### 2020-12-15 paper/paper.tex
draft of a draft

%
% File acl2020.tex
%
%% Based on the style files for ACL 2020, which were
%% Based on the style files for ACL 2018, NAACL 2018/19, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2020}
\usepackage{times}
\usepackage{latexsym}
\usepackage{amsfonts}
\renewcommand{\UrlFont}{\ttfamily\small}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B\textsc{ib}\TeX}

\title{Multi-Agent Reinforcement Learning in Zombie-Survival Games}

\author{Peter Hollows \\
  \texttt{stanford@dojo7.com} \\\And
  Mark Presser \\
  \texttt{mark.presser95@gmail.com} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  The abstract goes here...
\end{abstract}

\section{Introduction}

Reinforcement Learning (RL) methods are seeing increasing use in real-world control systems.
In part, the rising popularity of RL methods is due to its widely publicized success on challenging tasks,
as well as the availability of quality open-source frameworks containing implementations of state-of-the-art models.

\subsection{Closing the Reality Gap}

Current RL methods are not sample-effecicient, meaning the policy learner needs a lot of data before an effective policy can be found.
Often, there is not enough data available, or it cannot be captured safely (consider an autonomous driver learning the best policy for a near-crash scenario, where it will make mistakes in order to learn).
To work around these issues, RL policies intended for real-world control tasks are usually trained in simulated environments before being further trained or deployed in the real-world.

However, simulations often do not capture all the relevant real-world dynamics, so it can be difficult to transition a policy from the simulator to reality.
This difficulty, known as the "reality gap", is often mitigated by adding noise to the simulation (to mimic noisy sensors and stochastic dynamics),
or by sufficiently improving the simulation to model the relevant dynamics.

This approach has enjoyed success in single-agent RL settings like autonomous aircraft and robotics.

\subsection{Multi-Agent RL}

In cooperative settings like factories, or competitive settings like markets, using the simulation technique becomes more challenging;
simulating the dynamics of other agents present in the environment is a difficult problem, even in simulation.
This is especially the case when the policies of multiple agents are being modified and improved at the same time.

These issues in multi-agent settings motivated a training algorithm called Multi-Agent Deep Deterministic Policy Gradient (MADDPG).
MADDPG addresses some issues in multi-agent learning, and allows running experiments that explore the emergence of policies competitive and cooperative policies in complex environments.

\subsection{Environmental Dynamics}

We explore MADDPG in the context of a mixed competitive-cooperative simulated environment called Zombie-Survival game under multiple scenarios.
In these scenarios we show how specific kinds of environmental dynamics can affect large (and sometimes surprising and unintended) changes in the evolution of policies.
We present methods and results for each scenario, and conclude with a discussion (\ref{sec:discussion}) on safety issues we feel apply to a wide array of RL systems.

\subsection{Overview}

In (\ref{sec:background}) we review Markov Decision Processes (MDPs), the Actor-Critic algorithm, MADDPG, and agent design.
In (\ref{sec:game}) we detail the basic simulation environment.
In (\ref{sec:baseline}) we detail the experimental setup and initial results.
In (\ref{sec:anon}) we investigate partial observability, and how compression of the observation function can speed policy discovery.
In (\ref{sec:health}) we alter the environment to cause a change in the value function (keeping reward unchanged) to study the impact on learned policies.
In (\ref{sec:arms}) we arm the survivors with projectile weapons to investigate how additional capability can make it harder for agents to achieve their objectives.

\section{Background}
\label{sec:background}

\subsection{The Markov Decision Process}

A Markov Decision Process (MDP) formalizes the interaction of an agent with an environment.
At every time step, an \emph{agent} receives observations of the \emph{state} of the \emph{environment}, and sends \emph{actions} to the environment.
The environment transitions from one state to the next over time, and actions from the agent can influence which states the environment transitions to.
At every time step, the agent receives a \emph{reward} based on the state of the environment.
In the case when state transitions are not independent (i.e. the chance of arriving in a state depends on visiting previous states), then agent actions in a state can impact future expected rewards as well as immediate rewards.

RL algorithms attempt to solve the MDP in order to maximize return (reward over time).
This can be done by learning a policy (a function that maps from state to action).
A policy can often be learned directly, or by learning a value function (a mapping from states and actions to expected returns) from which a policy can be extracted.

\subsection{Actor-Critic methods}

Q-Learning is a popular RL method that makes use of an action-value function $Q^\pi(s, a)$ corresponding to a policy $\pi$ (where $s$ is the state, and $a$ the action).
Deep Q-Learning learns (DQN) learns the action-value function $Q^*$ corresponding to the optimal policy by minimizing the loss function
\begin{equation}
  \mathcal{L}(\theta)=\mathbb{E}_{s,a,r,s'}[(Q^*(s,a|\theta)-y)^2],
\end{equation}
where $y=r+\gamma\max_{a'}\bar{Q}^*(s',a')$,
by repeatedly adjusting the parameters $\theta$ of the target action-value function $\bar{Q}$.
The Q-function can be expressed as $Q^{\pi}(s,a)=\mathbb{E}[R|s,a]$, where $R_t$ is the discounted sum of future rewards at time $t$, given by: $R_t=\sum_{k=t}^\infty\gamma(k-t)r_k$.
To reduce the variance of estimates for the above expectation, Q-Learning often makes use of an Experience Replay (ER) buffer that stores the experiences of many state transitions.

In Actor-Critic (AC), both a policy (the actor) and an action-value function (the critic) are learned.
The Critic is used to estimate the policy gradient (the direction of changes to the policy parameters that cause actions that generate higher expected returns),
which is then used to update the policy function parameters at each training step.

\subsection{MADDPG}

The Markov Decision Process (MDP) framework regards other agents as part of the environment from the perspective of a single agent;
so if other agents in the environment are learning and modifying their behaviour then the environment contains shifting dynamics (the environmental data-generating process is considered non-stationary).
Deep Deterministic Policy Gradient (DDPG) is an actor-critic algorithm used in continuous control tasks.
It was shown in () that the chance of sampling a policy-gradient update in an improved direction decreases exponentially with the number of agents being trained in the shared environemnt.

MADDPG is a modified actor-critic method designed to address the stabilitiy issues due to the simultaneous learning of policies in the multi-agent setting.
A key component of MADDPG is the use of a shared critic that has access to the policies of all other agents during training.
In MADDPG, agents can have different state and action spaces, and be goverened by different reward functions.
We use the MADDPG algorithm for all agents within our scenarios.

\subsection{Agent design}

When implementing an RL system there are many important design considerations:

\begin{itemize}
  \item \emph{Algorithm}: Different algorithms are suited to different tasks. A selected algorithm will need to perform well with the selected action-space, state-space, reward function and environmental dynamics.
  \item \emph{State-space}: The state-space describes how environmental observations are presented to the algorithm. Constructing the correct, relevant abstractions inform how the agent `sees the world'.
  \item \emph{Action-space}: The action-space describes actions available to the agent. The level of abstraction here should be fine enough for accurate control, but general enough for the policy to be learnable.
  \item \emph{Reward function}: The reward function provides a scalar feedback value (positive or negative) to the agent. Most algorithms assume the reward signal contains no noise, and require point-estimates rather than a distribution, so crafting a consistent reward function that reflects the precise desires of the designer can be challenging.
\end{itemize}

We examine the imapct of design choices for state-space, action-space, and reward function in our scenarios.

\section{Scenarios}
\label{sec:scenarios}

\subsection{Zombie-Survival Game}
\label{sec:game}

The Zombie-Survival scenarios cover a set of games simulating survivors and zombies in a zombie-apocalypse scenario.
Survivors attempt to distance themselves from zombies and are penalized (-10 reward) for each contact.
Zombies attempt contact with survivors and are rewarded (+10 reward) for each contact.
Zombies have the advantage of numbers (there are 5 zombies and 2 survivors), whereas survivors have the advantage of faster accleration and a higher maximum speed.
All reward functions are independent, there are no explicitly shared rewards for outcomes.

The game offers a scenario with asymmetric power dynamics between competitive teams, and allows the emergence of compeitive or cooperative behavior within a team.

\subsection{Scenario: Baseline}
\label{sec:baseline}

We modify the 'Predator-prey' particle environment from (), which implements a multi-agent environment with a continuous action-space, a fully-observable state-space, and basic 2D physics.
Each agent is represented as a circle that has mass, position, velocity, velocity dampening, maximum acceleration, and maximum speed. Actions are 4D discrete (softmax probabilities) which are mapped to 2D continuous force vectors to control movement. The predator-prey environment uses reward shaping to discourage agents from moving out of bounds.
We replace the additional term in the reward function and implement fixed boundary walls.
We run several training jobs for 10,000 steps each to capture the emergence of different policies.

\subsection{Scenario: Anonymity}
\label{sec:anon}

\subsection{Scenario: Health}
\label{sec:health}

\subsection{Scenario: Armed}
\label{sec:arms}

\section{Discussion}
\label{sec:discussion}

FROM HERE FOLLOWS TEMPLATE STUFF

\begin{table}
\centering
\begin{tabular}{lrl}
\hline \textbf{Type of Text} & \textbf{Font Size} & \textbf{Style} \\ \hline
paper title & 15 pt & bold \\
author names & 12 pt & bold \\
author affiliation & 12 pt & \\
the word ``Abstract'' & 12 pt & bold \\
section titles & 12 pt & bold \\
subsection titles & 11 pt & bold \\
document text & 11 pt  &\\
captions & 10 pt & \\
abstract text & 10 pt & \\
bibliography & 10 pt & \\
footnotes & 9 pt & \\
\hline
\end{tabular}
\caption{\label{font-table} Font guide. }
\end{table}

\paragraph{\LaTeX-specific details:}
To use Times Roman in \LaTeX2e{}, put the following in the preamble:
\begin{quote}
\small
\begin{verbatim}
\usepackage{times}
\usepackage{latexsym}
\end{verbatim}
\end{quote}

\subsection{Ruler}
A printed ruler (line numbers in the left and right margins of the article) should be presented in the version submitted for review, so that reviewers may comment on particular lines in the paper without circumlocution.
The presence or absence of the ruler should not change the appearance of any other content on the page.
The camera ready copy should not contain a ruler.

\paragraph{Reviewers:}
note that the ruler measurements may not align well with lines in the paper -- this turns out to be very difficult to do well when the paper contains many figures and equations, and, when done, looks ugly.
In most cases one would expect that the approximate location will be adequate, although you can also use fractional references (\emph{e.g.}, this line ends at mark $295.5$).

\paragraph{\LaTeX-specific details:}
The style files will generate the ruler when {\small\verb|\aclfinalcopy|} is commented out, and remove it otherwise.

\subsection{Title and Authors}
\label{ssec:title-authors}

Center the title, author's name(s) and affiliation(s) across both columns.
Do not use footnotes for affiliations.
Place the title centered at the top of the first page, in a 15-point bold font.
Long titles should be typed on two lines without a blank line intervening.
Put the title 2.5 cm from the top of the page, followed by a blank line, then the author's names(s), and the affiliation on the following line.
Do not use only initials for given names (middle initials are allowed).
Do not format surnames in all capitals (\emph{e.g.}, use ``Mitchell'' not ``MITCHELL'').
Do not format title and section headings in all capitals except for proper names (such as ``BLEU'') that are
conventionally in all capitals.
The affiliation should contain the author's complete address, and if possible, an electronic mail address.

The title, author names and addresses should be completely identical to those entered to the electronical paper submission website in order to maintain the consistency of author information among all publications of the conference.
If they are different, the publication chairs may resolve the difference without consulting with you; so it is in your own interest to double-check that the information is consistent.

Start the body of the first page 7.5 cm from the top of the page.
\textbf{Even in the anonymous version of the paper, you should maintain space for names and addresses so that they will fit in the final (accepted) version.}

\subsection{Abstract}
Use two-column format when you begin the abstract.
Type the abstract at the beginning of the first column.
The width of the abstract text should be smaller than the
width of the columns for the text in the body of the paper by 0.6 cm on each side.
Center the word \textbf{Abstract} in a 12 point bold font above the body of the abstract.
The abstract should be a concise summary of the general thesis and conclusions of the paper.
It should be no longer than 200 words.
The abstract text should be in 10 point font.

\subsection{Text}
Begin typing the main body of the text immediately after the abstract, observing the two-column format as shown in the present document.

Indent 0.4 cm when starting a new paragraph.

\subsection{Sections}

Format section and subsection headings in the style shown on the present document.
Use numbered sections (Arabic numerals) to facilitate cross references.
Number subsections with the section number and the subsection number separated by a dot, in Arabic numerals.

\subsection{Footnotes}
Put footnotes at the bottom of the page and use 9 point font.
They may be numbered or referred to by asterisks or other symbols.\footnote{This is how a footnote should appear.}
Footnotes should be separated from the text by a line.\footnote{Note the line separating the footnotes from the text.}

\subsection{Graphics}

Place figures, tables, and photographs in the paper near where they are first discussed, rather than at the end, if possible.
Wide illustrations may run across both columns.
Color is allowed, but adhere to Section~\ref{ssec:accessibility}'s guidelines on accessibility.

\paragraph{Captions:}
Provide a caption for every illustration; number each one sequentially in the form:
``Figure 1. Caption of the Figure.''
``Table 1. Caption of the Table.''
Type the captions of the figures and tables below the body, using 10 point text.
Captions should be placed below illustrations.
Captions that are one line are centered (see Table~\ref{font-table}).
Captions longer than one line are left-aligned (see Table~\ref{tab:accents}).

\begin{table}
\centering
\begin{tabular}{lc}
\hline
\textbf{Command} & \textbf{Output}\\
\hline
\verb|{\"a}| & {\"a} \\
\verb|{\^e}| & {\^e} \\
\verb|{\`i}| & {\`i} \\ 
\verb|{\.I}| & {\.I} \\ 
\verb|{\o}| & {\o} \\
\verb|{\'u}| & {\'u}  \\ 
\verb|{\aa}| & {\aa}  \\\hline
\end{tabular}
\begin{tabular}{lc}
\hline
\textbf{Command} & \textbf{Output}\\
\hline
\verb|{\c c}| & {\c c} \\ 
\verb|{\u g}| & {\u g} \\ 
\verb|{\l}| & {\l} \\ 
\verb|{\~n}| & {\~n} \\ 
\verb|{\H o}| & {\H o} \\ 
\verb|{\v r}| & {\v r} \\ 
\verb|{\ss}| & {\ss} \\
\hline
\end{tabular}
\caption{Example commands for accented characters, to be used in, \emph{e.g.}, \BibTeX\ names.}\label{tab:accents}
\end{table}

\paragraph{\LaTeX-specific details:}
The style files are compatible with the caption and subcaption packages; do not add optional arguments.
\textbf{Do not override the default caption sizes.}

\subsection{Hyperlinks}
Within-document and external hyperlinks are indicated with Dark Blue text, Color Hex \#000099.

\subsection{Citations}
Citations within the text appear in parentheses as~\citep{Gusfield:97} or, if the author's name appears in the text itself, as \citet{Gusfield:97}.
Append lowercase letters to the year in cases of ambiguities.  
Treat double authors as in~\citep{Aho:72}, but write as in~\citep{Chandra:81} when more than two authors are involved. Collapse multiple citations as in~\citep{Gusfield:97,Aho:72}.

Refrain from using full citations as sentence constituents.
Instead of
\begin{quote}
  ``\citep{Gusfield:97} showed that ...''
\end{quote}
write
\begin{quote}
``\citet{Gusfield:97} showed that ...''
\end{quote}

\begin{table*}
\centering
\begin{tabular}{lll}
\hline
\textbf{Output} & \textbf{natbib command} & \textbf{Old ACL-style command}\\
\hline
\citep{Gusfield:97} & \small\verb|\citep| & \small\verb|\cite| \\
\citealp{Gusfield:97} & \small\verb|\citealp| & no equivalent \\
\citet{Gusfield:97} & \small\verb|\citet| & \small\verb|\newcite| \\
\citeyearpar{Gusfield:97} & \small\verb|\citeyearpar| & \small\verb|\shortcite| \\
\hline
\end{tabular}
\caption{\label{citation-guide}
Citation commands supported by the style file.
The style is based on the natbib package and supports all natbib citation commands.
It also supports commands defined in previous ACL style files for compatibility.
}
\end{table*}

\paragraph{\LaTeX-specific details:}
Table~\ref{citation-guide} shows the syntax supported by the style files.
We encourage you to use the natbib styles.
You can use the command {\small\verb|\citet|} (cite in text) to get ``author (year)'' citations as in \citet{Gusfield:97}.
You can use the command {\small\verb|\citep|} (cite in parentheses) to get ``(author, year)'' citations as in \citep{Gusfield:97}.
You can use the command {\small\verb|\citealp|} (alternative cite without  parentheses) to get ``author year'' citations (which is useful for  using citations within parentheses, as in \citealp{Gusfield:97}).

\subsection{References}
Gather the full set of references together under the heading \textbf{References}; place the section before any Appendices. 
Arrange the references alphabetically by first author, rather than by order of occurrence in the text.

Provide as complete a citation as possible, using a consistent format, such as the one for \emph{Computational Linguistics\/} or the one in the  \emph{Publication Manual of the American 
Psychological Association\/}~\citep{APA:83}.
Use full names for authors, not just initials.

Submissions should accurately reference prior and related work, including code and data.
If a piece of prior work appeared in multiple venues, the version that appeared in a refereed, archival venue should be referenced.
If multiple versions of a piece of prior work exist, the one used by the authors should be referenced.
Authors should not rely on automated citation indices to provide accurate references for prior and related work.

The following text cites various types of articles so that the references section of the present document will include them.
\begin{itemize}
\item Example article in journal: \citep{Ando2005}.
\item Example article in proceedings, with location: \citep{borschinger-johnson-2011-particle}.
\item Example article in proceedings, without location: \citep{andrew2007scalable}.
\item Example arxiv paper: \citep{rasooli-tetrault-2015}. 
\end{itemize}

\paragraph{\LaTeX-specific details:}
The \LaTeX{} and Bib\TeX{} style files provided roughly follow the American Psychological Association format.
If your own bib file is named \texttt{\small acl2020.bib}, then placing the following before any appendices in your \LaTeX{}  file will generate the references section for you:
\begin{quote}\small
\verb|\bibliographystyle{acl_natbib}|\\
\verb|\bibliography{acl2020}|
\end{quote}

You can obtain the complete ACL Anthology as a Bib\TeX\ file from \url{https://aclweb.org/anthology/anthology.bib.gz}.
To include both the anthology and your own bib file, use the following instead of the above.
\begin{quote}\small
\verb|\bibliographystyle{acl_natbib}|\\
\verb|\bibliography{anthology,acl2020}|
\end{quote}

\subsection{Digital Object Identifiers}
As part of our work to make ACL materials more widely used and cited outside of our discipline, ACL has registered as a CrossRef member, as a registrant of Digital Object Identifiers (DOIs), the standard for registering permanent URNs for referencing scholarly materials.

All camera-ready references are required to contain the appropriate DOIs (or as a second resort, the hyperlinked ACL Anthology Identifier) to all cited works.
Appropriate records should be found for most materials in the current ACL Anthology at \url{http://aclanthology.info/}.
As examples, we cite \citep{goodman-etal-2016-noise} to show you how papers with a DOI will appear in the bibliography.
We cite \citep{harper-2014-learning} to show how papers without a DOI but with an ACL Anthology Identifier will appear in the bibliography.

\paragraph{\LaTeX-specific details:}
Please ensure that you use Bib\TeX\ records that contain DOI or URLs for any of the ACL materials that you reference.
If the Bib\TeX{} file contains DOI fields, the paper title in the references section will appear as a hyperlink to the DOI, using the hyperref \LaTeX{} package.

\subsection{Appendices}
Appendices, if any, directly follow the text and the
references (but only in the camera-ready; see Appendix~\ref{sec:appendix}).
Letter them in sequence and provide an informative title:
\textbf{Appendix A. Title of Appendix}.

\section{Accessibility}
\label{ssec:accessibility}

In an effort to accommodate people who are color-blind (as well as those printing to paper), grayscale readability is strongly encouraged.
Color is not forbidden, but authors should ensure that tables and figures do not rely solely on color to convey critical distinctions.
A simple criterion:
All curves and points in your figures should be clearly distinguishable without color.

\section{Translation of non-English Terms}

It is also advised to supplement non-English characters and terms with appropriate transliterations and/or translations since not all readers understand all such characters and terms.
Inline transliteration or translation can be represented in the order of:
\begin{center}
\begin{tabular}{c}
original-form \\
transliteration \\
``translation''
\end{tabular}
\end{center}

\section{\LaTeX{} Compilation Issues}
You may encounter the following error during compilation: 
\begin{quote}
{\small\verb|\pdfendlink|} ended up in different nesting level than {\small\verb|\pdfstartlink|}.
\end{quote}
This happens when \texttt{\small pdflatex} is used and a citation splits across a page boundary.
To fix this, the style file contains a patch consisting of two lines:
(1) {\small\verb|\RequirePackage{etoolbox}|} (line 455 in \texttt{\small acl2020.sty}), and
(2) A long line below (line 456 in \texttt{\small acl2020.sty}).

If you still encounter compilation issues even with the patch enabled, disable the patch by commenting the two lines, and then disable the \texttt{\small hyperref} package by loading the style file with the \texttt{\small nohyperref} option:

\noindent
{\small\verb|\usepackage[nohyperref]{acl2020}|}

\noindent
Then recompile, find the problematic citation, and rewrite the sentence containing the citation. (See, {\em e.g.}, \url{http://tug.org/errors.html})

\section*{Acknowledgments}

The acknowledgments should go immediately before the references. Do not number the acknowledgments section.
Do not include this section when submitting your paper for review.

\bibliography{anthology,acl2020}
\bibliographystyle{acl_natbib}

\appendix

\section{Appendices}
\label{sec:appendix}
Appendices are material that can be read, and include lemmas, formulas, proofs, and tables that are not critical to the reading and understanding of the paper. 
Appendices should be \textbf{uploaded as supplementary material} when submitting the paper for review.
Upon acceptance, the appendices come after the references, as shown here.

\paragraph{\LaTeX-specific details:}
Use {\small\verb|\appendix|} before any appendix section to switch the section numbering over to letters.

\section{Supplemental Material}
\label{sec:supplemental}
Submissions may include non-readable supplementary material used in the work and described in the paper.
Any accompanying software and/or data should include licenses and documentation of research review as appropriate.
Supplementary material may report preprocessing decisions, model parameters, and other details necessary for the replication of the experiments reported in the paper.
Seemingly small preprocessing decisions can sometimes make a large difference in performance, so it is crucial to record such decisions to precisely characterize state-of-the-art methods.

Nonetheless, supplementary material should be supplementary (rather than central) to the paper.
\textbf{Submissions that misuse the supplementary material may be rejected without review.}
Supplementary material may include explanations or details of proofs or derivations that do not fit into the paper, lists of
features or feature templates, sample inputs and outputs for a system, pseudo-code or source code, and data.
(Source code and data should be separate uploads, rather than part of the paper).

The paper should not rely on the supplementary material: while the paper may refer to and cite the supplementary material and the supplementary material will be available to the reviewers, they will not be asked to review the supplementary material.

\end{document}

### 2020-12-16 experiments/train.py
revert debug loop

@@ -158,6 +158,7 @@ def train(arglist):
 
             # for displaying policies while training
             if arglist.display and (episode_number % arglist.display_rate == 0) and episode_number > 0 and er_fill_frac_min >= 1.0:
+            #if arglist.display and (episode_number % 5 == 0) and episode_number > 0:
                 time.sleep(0.1)
                 env.render()
 

### 2020-12-16 multiagent/scenarios/zombie_arms.py
zombie_arms: pistols with no external reload observation

@@ -42,11 +42,11 @@ class Scenario(BaseScenario):
 
             # weapons
             agent.armed = True
-            agent.arms_reload_time = 1.0
-            agent.arms_pallet_count = 10
-            agent.arms_pallet_damage = 0.1
-            agent.arms_pallet_range = 2
-            agent.arms_pallet_spread = 5 /360.0*2*np.pi
+            agent.arms_reload_time = 0.2
+            agent.arms_pallet_count = 2
+            agent.arms_pallet_damage = 0.2
+            agent.arms_pallet_range = 3
+            agent.arms_pallet_spread = 3 /360.0*2*np.pi
 
             agents.append(agent)
         # add zombies' team
@@ -116,15 +116,16 @@ class Scenario(BaseScenario):
         #   vel         2
         #   ang         1
         #   health      1
+        #   reloading   1
         #
         # others: (sorted by team then distance, must be constant size within team)
         #
         #   rel_pos     2
         #   rel_vel     2
         #   rel_ang     1
-        #   reloading   1
 
-        s = 2+2+1 + (2+2+1+1)*len(other_agents)
+        s = 2+2+1 + (2+2+1)*len(other_agents)
+        if agent.armed: s += 1
         obs = np.zeros((s,))
         i = 0
 
@@ -132,6 +133,8 @@ class Scenario(BaseScenario):
         obs[i:i+2] = agent.state.p_pos;     i += 2
         obs[i:i+2] = agent.state.p_vel;     i += 2
         obs[i:i+1] = agent.state.health;    i += 1
+        if agent.armed:
+            obs[i:i+1] = agent.state.reloading; i += 1
 
         abs_ang = agent.state.aim_heading if agent.armed else 0.
         for other in other_agents:
@@ -143,9 +146,6 @@ class Scenario(BaseScenario):
             # relative angle to current heading (not relative heading)
             rel_ang = np.arctan2(rel_pos[1], rel_pos[0]) - abs_ang
             obs[i:i+1] = (rel_ang % 2*np.pi) - np.pi; i += 1
-            # reloading
-            rel = other.state.reloading if other.armed else 0.
-            obs[i:i+1] = rel; i += 1
 
         assert i == s
 

### 2020-12-16 paper/paper.tex
update paper

@@
 \usepackage{amsfonts}
+\usepackage{amsmath}
 \renewcommand{\UrlFont}{\ttfamily\small}
@@
 
-In (\ref{sec:background}) we review Markov Decision Processes (MDPs), the Actor-Critic algorithm, MADDPG, and agent design.
+In (\ref{sec:background}) we review theory, RL algorithms, and aspects of agent design.
 In (\ref{sec:game}) we detail the basic simulation environment.
@@
 
-\subsection{Actor-Critic methods}
+\subsection{DDPG}
 
@@
 
-In Actor-Critic (AC), both a policy (the actor) and an action-value function (the critic) are learned.
-The Critic is used to estimate the policy gradient (the direction of changes to the policy parameters that cause actions that generate higher expected returns),
-which is then used to update the policy function parameters at each training step.
+Policy-gradient (PG) methods learn a policy $\pi_\theta$, parameterized by $\theta$,
+by sampling the gradient of a performance measure $J(\theta)$ with respect to $\theta$.
+
+\begin{equation}
+  \nabla_\theta J(\theta)=\mathbb{E}_\pi\left[Q^\pi(s, a) \frac{\nabla_\theta\pi_\theta(a|s)}{\pi_\theta(a|s)}\right]
+\end{equation}
+
+REINFORCE, a simple PG algorithm, uses all the steps in an episode to find $Q^\pi(s,a)=\mathbb{E}_\pi[G|s,a]$, where $G$ is the total reward of an episode.
+Because reward functions sometimes give high values for some states, this method can give high variance gradient estimates.
+This can be remedied by using a baseline $b(s)$:
+
+\begin{equation}
+  \nabla_\theta J(\theta)=\mathbb{E}_\pi\left[(G-b(s)) \frac{\nabla_\theta\pi_\theta(a|s)}{\pi_\theta(a|s)}\right]
+\end{equation}
+
+Actor-Critic (AC) improves on REINFORCE by learning a state-value function with baseline (critic), and an action-value function (actor) concurrently.
+AC also has the advantage that the parameters for both functions are updated at every step in an episode.
+The policy gradient at each time step $t$ can be expressed:
+
+\begin{equation}
+  \nabla_\theta J(\theta)=\mathbb{E}_\pi\left[\delta_t \frac{\nabla_\theta\pi_\theta(a|s)}{\pi_\theta(a|s)}\right]
+\end{equation}
+\begin{equation}
+  \delta_t = r_{t+1}+\gamma \hat{v}(s_{t+1}, w)-\hat{v}(s_t,w)
+\end{equation}
+
+where $\gamma$ is a discount parameter, and $\hat{v}(w)$ the state-value function with parameters $w$.
+
+Deterministic Policy Gradient (DPG) methods extend policy gradient to deterministic policies.
+A deterministic policy ($\mu_\theta$) maps a state to an action,
+whereas a stochastic policy ($\pi_\theta$) provides a conditional probability of an action in a given state.
+
+\begin{equation}
+  \nabla_\theta J(\theta)=\mathbb{E}_s\left[\nabla_\theta \mu_\theta(a|s) \nabla_a Q^\mu(s,a)|_{a=\mu_\theta(s)}\right]
+\end{equation}
+
+DPG methods are often used in continuous control tasks
+(taking the gradient of $Q^\mu$ with respect to $a$ requires the action space to be continuous).
+
+Deep Deterministic Policy Gradient (DDPG) combines ideas from DPG and DQN.
+In DDPG the actor and critic are approximated with neural networks.
+DDPG uses target networks for both actor and critic, and makes use of an experience replay buffer to store agent trajectories.
 
@@
 
-The Markov Decision Process (MDP) framework regards other agents as part of the environment from the perspective of a single agent;
-so if other agents in the environment are learning and modifying their behaviour then the environment contains shifting dynamics (the environmental data-generating process is considered non-stationary).
-Deep Deterministic Policy Gradient (DDPG) is an actor-critic algorithm used in continuous control tasks.
+The Markov Decision Process (MDP) framework regards other agents as part of the environment from the perspective of a single agent.
+From the perspective of an agent, the influence on rewards from other agents in the environment can make it harder to learn a policy;
+this is especially so when the environment contains other agents that are also learning policies,
+because elements of the `environment' change while the agent is trying to learn them.
+
 It was shown in () that the chance of sampling a policy-gradient update in an improved direction decreases exponentially with the number of agents being trained in the shared environemnt.
 
-MADDPG is a modified actor-critic method designed to address the stabilitiy issues due to the simultaneous learning of policies in the multi-agent setting.
-A key component of MADDPG is the use of a shared critic that has access to the policies of all other agents during training.
-In MADDPG, agents can have different state and action spaces, and be goverened by different reward functions.
+MADDPG addresses this stabilitiy issue by using centralized critics.
+In MADDPG, the critic $Q^\pi_i(x, a_1, a_2, ..., a_N)$ for agent $i$ has access to the actions of all $N$ agents, and some state information $x$.
+The critic is often only used in training, so it's possible to deploy MADDPG actors in continuous control tasks where knowledge of other actors' actions are not available.
+Agents can have different state and action spaces, and be goverened by different reward functions.
 We use the MADDPG algorithm for all agents within our scenarios.

### 2020-12-17 paper/paper.tex
paper update

@@
   \item \emph{State-space}: The state-space describes how environmental observations are presented to the algorithm.
-Constructing the correct, relevant abstractions inform how the agent `sees the world'.
+Constructing the correct, relevant abstractions allow an agent to better learn the environment.
   \item \emph{Action-space}: The action-space describes actions available to the agent.
-The level of abstraction here should be fine enough for accurate control, but general enough for the policy to be learnable.
+The level of abstraction here should be fine enough for precise control, but general enough for a policy to be learnable.
   \item \emph{Reward function}: The reward function provides a scalar feedback value (positive or negative) to the agent.
-Most algorithms assume the reward signal contains no noise, and require point-estimates rather than a distribution, so crafting a consistent reward function that reflects the precise desires of the designer can be challenging.
+Most algorithms assume the reward signal contains no noise, and require point-estimates rather than a distribution.
+Crafting a consistent reward function that reflects the exact desires of the designer can be challenging.
+Some algorithms can infer the a reward function from feedback, but most will try to maximize explicit reward.
 \end{itemize}
@@
 The Zombie-Survival scenarios cover a set of games simulating survivors and zombies in a zombie-apocalypse scenario.
-Survivors attempt to distance themselves from zombies and are penalized (-10 reward) for each contact.
-Zombies attempt contact with survivors and are rewarded (+10 reward) for each contact.
+When a zombie is in physical contact with a survivor this results in a `bite' between zombie and survivor.
+Zombies attempt to bite survivors and are rewarded for each step they are in contact.
+Survivors attempt to distance themselves from zombies and are penalized every time they are bitten.
+Survivors can suffer a bite from multiple zombies in the same time step.
+
 Zombies have the advantage of numbers (there are 5 zombies and 2 survivors), whereas survivors have the advantage of faster accleration and a higher maximum speed.
-All reward functions are independent, there are no explicitly shared rewards for outcomes.
+All reward functions are independent, agents do not share reward.
+The game offers a scenario with asymmetric power dynamics between competitive teams, and allows the emergence of compeitive or cooperative behavior within a team.
 
-The game offers a scenario with asymmetric power dynamics between competitive teams, and allows the emergence of compeitive or cooperative behavior within a team.
+\begin{table}
+\centering
+\begin{tabular}{lrr}
+\hline \textbf{Setting} & \textbf{Survivors} & \textbf{Zombies} \\ \hline
+  world size ($m$) & 2x2 & 2x2 \\
+  timestep ($s$) & 0.1 & 0.1 \\
+  episode length ($s$) & 10 & 10 \\
+  num. agents & 2 & \textbf{5} \\
+  agent radius ($m$) & 0.05 & 0.05 \\
+  agent mass ($kg$) & 1.0 & 1.0 \\
+  acceleration ($ms^-2$) & \textbf{4.0} & 3.0 \\
+  max. speed ($ms^-1$) & \textbf{1.3} & 1.0 \\
+  bite reward & -10 & +10 \\
+\hline
+\end{tabular}
+\caption{\label{game-settings} Scenario settings. }
+\end{table}
 
@@
 Each agent is represented as a circle that has mass, position, velocity, velocity dampening, maximum acceleration, and maximum speed.
-Actions are 4D discrete (softmax probabilities) which are mapped to 2D continuous force vectors to control movement.
+%Actions are 4D discrete (softmax probabilities) which are mapped to 2D continuous force vectors to control movement.
 The predator-prey environment uses reward shaping to discourage agents from moving out of bounds.
-We replace the additional term in the reward function and implement fixed boundary walls.
+We remove this additional term in the reward function and implement fixed boundary walls.
+We implement the biting rules, modify the agent count and adjust the agents' physical capabilities (see Table \ref{game-settings}).
 We run several training jobs for 10,000 steps each to capture the emergence of different policies.
+
+\subsection{Results: Baseline}
 
@@
 
+In the \emph{anonymity} scenario we modify the \emph{baseline} scenario to introduce partial observability.
+We re-order agents' state vectors each step so that agents within teams are ordered by distance to the observing agent.
+To a survivor, a zombie now appears the same as any other zombie, and does not have identity by virtue of index in the state vector.
+We expect both survivors and zombies to learn more general policies with faster convergence than the baseline.
+
+\subsection{Results: Anonymity}
+
 \subsection{Scenario: Health}
@@
 
+In the \emph{health} scenario we modify the \emph{anonymity} scenario to introduce a health mechanism for survivors.
+Survivors start each episode with 100\% health, which for survivors is discounted (by 0.99) every bite.
+An agent's health level is added to the private observation space for the agent.
+Reward remains the same for bites regardless of health levels,
+however maximum speed at each step is limited to the fraction of health remaining.
+
+If a survivor is bitten too often, they become slower than the zombies and risk being overrun, leading to a large amount of bites for the remainder of the episode.
+This scenario does not alter the reward function, but increases the variance and magnitude of state-value estimates for both zombies and survivors.
+Additionally, this scenario shifts the balance of power heavily toward the zombies.
+We expect to see survivor policies emerge that are more risk-averse, and to see zombies elect to chase slower-moving survivors.
+
+\subsection{Results: Health}
+
 \subsection{Scenario: Armed}
 \label{sec:arms}
+
+In the \emph{armed} scenario we modify the \emph{health} scenario by arming the survivors with pistols.
+Each armed agent has a pistol that fires 2 shots when activated.
+On firing, two rays are cast with a small random spread ($\angle p \sim \mathcal{N}(\angle aim, 2^\circ)$) for a range (2m) that covers most of the area.
+When a ray first intersects with another agent it is counted as a `hit'.
+An agent experiencing a hit has their health level reduced (by 0.2).
+Armed agents have a reload time (1 step).
+Both zombies and survivors can be rendered immobile through reduced health by being hit too often.
+
+We add a normalized current heading and relative heading of locations of other agents to an armed agent's state space, as well as a normalized time remaining until reloaded.
+Reloading is automatic and does not require an action.
+The continuous Aiming action (left and right) modifies angular velocity of aim directly up to a maximum speed of 1 revolution per second.
+The firing mechanic is implemented as a `tirgger pressure' which determines the probability of a loaded pistol firing in a given timestep.
+Firing probability is implemented as the sigmoid function $p(fire) = \frac{1}{1+e^{20(x-0.5)}}$, where $x$ is the pressure on the trigger.
+
+We expect introducing the arms mechanic will move the balance of power back toward the survivors.
+We're not sure if the social dilemma presented will cause the survivors to fire on one another more often than the zombies,
+but make no adjustment to the reward function to discourage it.
+
+\subsection{Results: Armed}
 
@@
 FROM HERE FOLLOWS TEMPLATE STUFF
-
-\begin{table}
-\centering
-\begin{tabular}{lrl}
-\hline \textbf{Type of Text} & \textbf{Font Size} & \textbf{Style} \\ \hline
-paper title & 15 pt & bold \\
-author names & 12 pt & bold \\
-author affiliation & 12 pt & \\
-the word ``Abstract'' & 12 pt & bold \\
-section titles & 12 pt & bold \\
-subsection titles & 11 pt & bold \\
-document text & 11 pt  &\\
-captions & 10 pt & \\
-abstract text & 10 pt & \\
-bibliography & 10 pt & \\
-footnotes & 9 pt & \\
-\hline
-\end{tabular}
-\caption{\label{font-table} Font guide. }
-\end{table}
 

### 2020-12-18 paper/paper.tex
update paper

@@
-%
 % File acl2020.tex
@@
 \usepackage{amsmath}
+\usepackage{graphicx}
+
 \renewcommand{\UrlFont}{\ttfamily\small}
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{figures/baseline_reward.png}
+  \caption{Reward (moving average over 100 episodes) for survivors (green) and zombies (red) for 10 training runs.}
+  \label{fig:baseline_reward}
+\end{figure}
+
+There was some variability between training runs in the baseline scenario.
+Figure \ref{fig:baseline_reward} shows agent reward graphs over multiple runs.
+The reward trace for $r_5$ shows the zombie agent experienced a spike in reward corresponding to a drop in reward for a survivor in $r_0$
+that occurred only once in 10 training runs.
+Most agents show a similar pattern for at least one run.
+
+On further investigation we discovered the reason for this; a zombie usually fixates on a target, but zombies learn policies at different rates.
+Figure \ref{fig:baseline_speed} shows speed and reward over time for a single training run.
+We see that one survivor is not pursued as often to start with, but then recieves increasing attention from zombies learning to persure it.
+Following this attention the survivor learns to move much faster.
+We observed similar patterns in other runs.
+
+In the baseline state-space, every agent is unique, so all agents have to learn their expected reward relating to each other.
+From an agent's perspective, when considering reward the most important thing about an agent is whether they are a survivor or a zombie, not necessarily which survivor or zombie.
+We'd prefer agents learn general policies that consider other agents exchangeable within a team.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{figures/baseline_speed.png}
+  \caption{Moving averages for reward and speed in baseline scenario. Survivors in blue and orange.}
+  \label{fig:baseline_speed}
+\end{figure}
+
 \subsection{Scenario: Anonymity}
@@
 To a survivor, a zombie now appears the same as any other zombie, and does not have identity by virtue of index in the state vector.
+This reduces the size of the state-space considerably, and allows agents to learn policies that focus on nearby agents.
+
 We expect both survivors and zombies to learn more general policies with faster convergence than the baseline.
+We also expect to see zombies chasing which ever survivor is easiest to chase, rather than fixating on a single survivor.
 

### 2020-12-18 paper/paper.tex
paper update

@@
 Each agent is represented as a circle that has mass, position, velocity, velocity dampening, maximum acceleration, and maximum speed.
-%Actions are 4D discrete (softmax probabilities) which are mapped to 2D continuous force vectors to control movement.
 The predator-prey environment uses reward shaping to discourage agents from moving out of bounds.
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{figures/baseline_speed.png}
+  \caption{Reward and speed (moving averages over 200 episodes) for a single training run in baseline scenario. Survivors in blue and orange.}
+  \label{fig:baseline_speed}
+\end{figure}
+
 On further investigation we discovered the reason for this; a zombie usually fixates on a target, but zombies learn policies at different rates.
@@
 We'd prefer agents learn general policies that consider other agents exchangeable within a team.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.5\textwidth]{figures/baseline_speed.png}
-  \caption{Moving averages for reward and speed in baseline scenario. Survivors in blue and orange.}
-  \label{fig:baseline_speed}
-\end{figure}
 
@@
 
+We observed faster convergence to a stable policy (as judged by average speed and reward) when compared to baseline (see Figure \ref{fig:anon_base_compare}).
+We also observed more uniform persuit and biting (see Figure \ref{fig:anon_base_dist_bite}).
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{figures/anon_base_compare.png}
+  \caption{
+    Speed and reward evolution for baseline and anonymous scenarios.
+    Results taken as moving averages over 200 episodes prior to aggregation.
+  }
+  \label{fig:anon_base_compare}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{figures/anon_base_dist_bite.png}
+  \caption{
+    Distance and bite count comparison between baseline and anonymous scenarios.
+    Plot colors indicate average distance and average bite count per episode average across the last 5000 episodes of a training run.
+    Both axes are agent indexes, and color indicates relative value with white being the highest value.
+  }
+  \label{fig:anon_base_dist_bite}
+\end{figure}
+
 \subsection{Scenario: Health}
@@
 In the \emph{health} scenario we modify the \emph{anonymity} scenario to introduce a health mechanism for survivors.
-Survivors start each episode with 100\% health, which for survivors is discounted (by 0.99) every bite.
+Survivors start each episode with 100\% health, which for survivors is discounted ($\gamma=0.99$) every bite.
 An agent's health level is added to the private observation space for the agent.
@@
 however maximum speed at each step is limited to the fraction of health remaining.
+If a survivor is bitten too often, they become slower than the zombies and risk being overrun, leading to a large amount of bites for the remainder of the episode.
 
-If a survivor is bitten too often, they become slower than the zombies and risk being overrun, leading to a large amount of bites for the remainder of the episode.
 This scenario does not alter the reward function, but increases the variance and magnitude of state-value estimates for both zombies and survivors.
@@
 \subsection{Results: Health}
+
+Compared to the \emph{anonymous} scenario, agents in the \emph{health} scenario were slower on average (see Figure \ref{fig:anon_health_compare}, plot 1).
+It's interesting to note this slowdown occurred before the reward ramp-up due to zombies successfully immobilizing survivors.
+In part the slower speeds can be explained through partial health reductions, but rendering the simulation frames provided another explanation.
+In previous scenarios, survivors would run into walls at maximum speed to bounce back through the persuring group of zombies without losing speed.
+The bouncing strategy sacrifices minimal reward and allows a survivor to keep their distance ahead of the pursuing zombies after the manouver.
+In the \emph{health} scenario, this strategy becomes unsustainable, as every contact with a zombie (during the bounce) reduces the survivors maximum speed
+leading to a very bad outcome after only a few bounces.
+
+With health to consider, survivors opted for more impressive fake-and-dodge manouvers to avoid contact altogether.
+Often this would involve pausing to allow zombies to move closer to their target, then making use of the survivors' higher acceleration to escape at an angle orthagonal to the zombie group's velocity.
+
+However, clever dodging only worked in some instances.
+We measured a drastic increase in average zombie reward (see Figure \ref{fig:anon_health_compare}, plot 2) due to zombies learning to immobilize and bite survivors.
+In most cases, one survivor who would succumb to the zombies first, attracting the attention of most of the zombies and allowing the remaining survivor to stay out of harm's way.
+
+We found it difficult to predict at the beginning of an episode which survivor would ultimately attract the zombies.
+In watching many episodes we found a survivor having many zombies in tow would move directly toward the other survivor (who was often be keeping their distance in a corner).
+The first survivor would have higher velocity, and use it to risk a wall-bounce manouver near the second slower-moving one.
+The first survivor in this instance would retain the speed required to escape while the zombies switched their persuit to overrun the second survivor.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{figures/anon_health_compare.png}
+  \caption{
+    Comparison between anonymous and health scenario policy evolution.
+    Results taken as moving averages over 200 episodes prior to aggregation.
+    Plot 1: average speed of all agents.
+    Plot 2: average reward of zombies.
+    Plot 3: lowest health of all survivors at the conclusion of an episode.
+    Plot 4: the difference between survivor health at the conclusion of an episode.
+  }
+  \label{fig:anon_health_compare}
+\end{figure}
 
@@
 An agent experiencing a hit has their health level reduced (by 0.2).
-Armed agents have a reload time (1 step).
+Armed agents require time to reload (1 step).
 Both zombies and survivors can be rendered immobile through reduced health by being hit too often.

### 2020-12-18 paper/paper.tex
update paper

@@
 % File acl2020.tex
-%
-%% Based on the style files for ACL 2020, which were
-%% Based on the style files for ACL 2018, NAACL 2018/19, which were
-%% Based on the style files for ACL-2015, with some improvements
-%%  taken from the NAACL-2016 style
-%% Based on the style files for ACL-2014, which were, in turn,
-%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
-%% EACL-2009, IJCNLP-2008...
-%% Based on the style files for EACL 2006 by 
-%%e.agirre@ehu.es or Sergi.Balari@uab.es
-%% and that of ACL 08 by Joakim Nivre and Noah Smith
+%% Based on the style files for ACL 2020
 
@@
 
-\author{Peter Hollows \\
-  \texttt{stanford@dojo7.com} \\\And
-  Mark Presser \\
-  \texttt{mark.presser95@gmail.com} \\}
+\author{
+  Peter Hollows
+  \thanks {
+    Group submission for the final paper of XCS229ii,
+    \mbox{Stanford Center for Professional Development}
+  }
+  \\ \texttt{stanford@dojo7.com} \\
+  \And
+  Mark Presser
+  \footnotemark[1]
+  \\ \texttt{mark.presser95@gmail.com} \\
+}
 
@@
 However, simulations often do not capture all the relevant real-world dynamics, so it can be difficult to transition a policy from the simulator to reality.
-This difficulty, known as the "reality gap", is often mitigated by adding noise to the simulation (to mimic noisy sensors and stochastic dynamics),
+This difficulty, known as the ``reality gap'', is often mitigated by adding noise to the simulation (to mimic noisy sensors and stochastic dynamics),
 or by sufficiently improving the simulation to model the relevant dynamics.
@@
 
-These issues in multi-agent settings motivated a training algorithm called Multi-Agent Deep Deterministic Policy Gradient (MADDPG).
-MADDPG addresses some issues in multi-agent learning, and allows running experiments that explore the emergence of policies competitive and cooperative policies in complex environments.
-
-\subsection{Environmental Dynamics}
-
-We explore MADDPG in the context of a mixed competitive-cooperative simulated environment called Zombie-Survival game under multiple scenarios.
-In these scenarios we show how specific kinds of environmental dynamics can affect large (and sometimes surprising and unintended) changes in the evolution of policies.
-We present methods and results for each scenario, and conclude with a discussion (\ref{sec:discussion}) on safety issues we feel apply to a wide array of RL systems.
+These issues in multi-agent settings motivated a training algorithm called Multi-Agent Deep Deterministic Policy Gradient (MADDPG) \citep{lowe2020multiagent}.
+MADDPG addresses some issues in multi-agent learning, and allows running experiments that explore the emergence of competitive and cooperative policies in complex environments.
 
 \subsection{Overview}
+
+We explore MADDPG in the context of a mixed competitive-cooperative simulated environment called ``Zombie-Survival''.
+We make several modifications in order to test different scenarios.
+In these scenarios we show how specific kinds of environmental dynamics can affect changes in the evolution of policies.
 
@@
 In (\ref{sec:baseline}) we detail the experimental setup and initial results.
-In (\ref{sec:anon}) we investigate partial observability, and how compression of the observation function can speed policy discovery.
-In (\ref{sec:health}) we alter the environment to cause a change in the value function (keeping reward unchanged) to study the impact on learned policies.
+In (\ref{sec:anon}) we investigate partial observability, and how compression of the observation-space can speed policy discovery.
+In (\ref{sec:health}) we alter the environment to cause a change in the value-function (keeping reward unchanged) to study the impact on learned policies.
 In (\ref{sec:arms}) we arm the survivors with projectile weapons to investigate how additional capability can make it harder for agents to achieve their objectives.
+We present methods and results for each scenario, and conclude with a discussion (\ref{sec:discussion}) of our findings.
 
@@
 
-\subsection{DDPG}
+\subsection{Algorithms}
 
-Q-Learning is a popular RL method that makes use of an action-value function $Q^\pi(s, a)$ corresponding to a policy $\pi$ (where $s$ is the state, and $a$ the action).
-Deep Q-Learning learns (DQN) learns the action-value function $Q^*$ corresponding to the optimal policy by minimizing the loss function
+\paragraph{Deep Q-Learning Networks:} Q-Learning \citep{watkins} is an RL method that makes use of an action-value function $Q^\pi(s, a)$ corresponding to a policy $\pi$ (where $s$ is the state, and $a$ the action).
+Deep Q-Learning Networks (DQN) \citep{2015Natur.518..529M} learn the action-value function $Q^*$ corresponding to the optimal policy by minimizing the loss function
 \begin{equation}
@@
 The Q-function can be expressed as $Q^{\pi}(s,a)=\mathbb{E}[R|s,a]$, where $R_t$ is the discounted sum of future rewards at time $t$, given by: $R_t=\sum_{k=t}^\infty\gamma(k-t)r_k$.
-To reduce the variance of estimates for the above expectation, Q-Learning often makes use of an Experience Replay (ER) buffer that stores the experiences of many state transitions.
+To reduce the variance of estimates for the above expectation, DQN makes use of an Experience Replay (ER) buffer that stores the experiences of many state transitions.
 
-Policy-gradient (PG) methods learn a policy $\pi_\theta$, parameterized by $\theta$,
+\paragraph{Policy-gradient methods} learn a policy $\pi_\theta$, parameterized by $\theta$,
 by sampling the gradient of a performance measure $J(\theta)$ with respect to $\theta$.
@@
 
-REINFORCE, a simple PG algorithm, uses all the steps in an episode to find $Q^\pi(s,a)=\mathbb{E}_\pi[G|s,a]$, where $G$ is the total reward of an episode.
-Because reward functions sometimes give high values for some states, this method can give high variance gradient estimates.
-This can be remedied by using a baseline $b(s)$:
+\paragraph{REINFORCE} \citep{Williams92simplestatistical}, a simple policy-gradient algorithm, uses all the steps in an episode to find $Q^\pi(s,a)=\mathbb{E}_\pi[G|s,a]$, where $G$ is the total reward of an episode.
+Because reward functions sometimes give high values for some states, this method can give high-variance gradient estimates.
+This can be remedied by using a baseline $b(s)$ \citep{SuttonBarto}:
 
@@
 
-Actor-Critic (AC) improves on REINFORCE by learning a state-value function with baseline (critic), and an action-value function (actor) concurrently.
-AC also has the advantage that the parameters for both functions are updated at every step in an episode.
+\paragraph{Actor-critic} \citep{bhatnagar2009natural} improves on REINFORCE by learning a state-value function with baseline (critic), and an action-value function (actor) concurrently.
+Actor-critic also has the advantage that the parameters for both functions are updated at every step in an episode.
 The policy gradient at each time step $t$ can be expressed:
@@
 
-Deterministic Policy Gradient (DPG) methods extend policy gradient to deterministic policies.
+\paragraph{Deterministic Policy Gradient (DPG)} \citep{pmlr-v32-silver14} algorithms extend policy gradient methods to deterministic policies.
 A deterministic policy ($\mu_\theta$) maps a state to an action,
@@
 
-Deep Deterministic Policy Gradient (DDPG) combines ideas from DPG and DQN.
+\paragraph{Deep Deterministic Policy Gradient (DDPG)} \citep{lillicrap2019continuous} combines ideas from DPG and DQN.
 In DDPG the actor and critic are approximated with neural networks.
@@
 
-It was shown in () that the chance of sampling a policy-gradient update in an improved direction decreases exponentially with the number of agents being trained in the shared environemnt.
+\citealp{lowe2020multiagent} showed that the chance of sampling a policy-gradient update in an improved direction decreases exponentially with the number of agents being trained in the shared environemnt.
 
@@
 
-\subsection{Agent design}
+\subsection{Agent Design}
 
@@
   \item \emph{Reward function}: The reward function provides a scalar feedback value (positive or negative) to the agent.
-Most algorithms assume the reward signal contains no noise, and require point-estimates rather than a distribution.
-Crafting a consistent reward function that reflects the exact desires of the designer can be challenging.
-Some algorithms can infer the a reward function from feedback, but most will try to maximize explicit reward.
+Some algorithms can model uncertainty in the reward function \citep{ghavamzadeh2016bayesian}, but most expect a point-estimate rather than a distribution.
+Crafting a consistent reward function that reflects the exact desires of the designer can be challenging \citep{leike2017ai}.
+Some algorithms can infer the a reward function from feedback \citep{ng2000algorithms}, but most simply try to maximize the explicit reward from a hard-coded function.
 \end{itemize}
 
-We examine the imapct of design choices for state-space, action-space, and reward function in our scenarios.
+In our scenarios we examine the imapct of design choices for state-space and action-space under a fixed reward-function.
 
@@
 All reward functions are independent, agents do not share reward.
-The game offers a scenario with asymmetric power dynamics between competitive teams, and allows the emergence of compeitive or cooperative behavior within a team.
+The game offers a scenario with asymmetric power dynamics between competitive teams, and allows the emergence of competitive or cooperative behavior within a team.
 
@@
 
-We modify the 'Predator-prey' particle environment from (), which implements a multi-agent environment with a continuous action-space, a fully-observable state-space, and basic 2D physics.
+We modify the ``Predator-prey'' particle environment from \citealp{lowe2020multiagent}, which implements a multi-agent environment with a continuous action-space, a fully-observable state-space, and basic 2D physics.
 Each agent is represented as a circle that has mass, position, velocity, velocity dampening, maximum acceleration, and maximum speed.
@@
 
-There was some variability between training runs in the baseline scenario.
+\paragraph{Variability:} We observed variability in policy convergence between training runs.
 Figure \ref{fig:baseline_reward} shows agent reward graphs over multiple runs.
@@
 
-On further investigation we discovered the reason for this; a zombie usually fixates on a target, but zombies learn policies at different rates.
+\paragraph{Target fixation:} On further investigation we discovered the reason for the variability; a zombie usually fixates on a target, but zombies learn policies at different rates.
 Figure \ref{fig:baseline_speed} shows speed and reward over time for a single training run.
@@
 
-Compared to the \emph{anonymous} scenario, agents in the \emph{health} scenario were slower on average (see Figure \ref{fig:anon_health_compare}, plot 1).
+\paragraph{Breakout:} Compared to the \emph{anonymous} scenario, agents in the \emph{health} scenario were slower on average (see Figure \ref{fig:anon_health_compare}, plot 1).
 It's interesting to note this slowdown occurred before the reward ramp-up due to zombies successfully immobilizing survivors.
@@
 
-However, clever dodging only worked in some instances.
-We measured a drastic increase in average zombie reward (see Figure \ref{fig:anon_health_compare}, plot 2) due to zombies learning to immobilize and bite survivors.
-In most cases, one survivor who would succumb to the zombies first, attracting the attention of most of the zombies and allowing the remaining survivor to stay out of harm's way.
+\paragraph{Game-over:} We measured a drastic increase in average zombie reward (see Figure \ref{fig:anon_health_compare}, plot 2) due to zombies learning to immobilize and bite survivors.
+In most cases, one survivor would succumb to the zombies first.
+Zombies learn that moving in the direction of the slower-moving survivor leads to states with higher expected value, so they collaborate in slowing down one survivor, leaving the other survivor to escape out of harm's way.
 
-We found it difficult to predict at the beginning of an episode which survivor would ultimately attract the zombies.
-In watching many episodes we found a survivor having many zombies in tow would move directly toward the other survivor (who was often be keeping their distance in a corner).
-The first survivor would have higher velocity, and use it to risk a wall-bounce manouver near the second slower-moving one.
-The first survivor in this instance would retain the speed required to escape while the zombies switched their persuit to overrun the second survivor.
+%We found it difficult to predict at the beginning of an episode which survivor would ultimately attract the zombies.
+%In watching many episodes we found a survivor having many zombies in tow would move directly toward the other survivor (who was often be keeping their distance in a corner).
+%The first survivor would have higher velocity, and use it to risk a wall-bounce manouver near the second slower-moving one.
+%The first survivor in this instance would retain the speed required to escape while the zombies switched their persuit to overrun the second survivor.
 
@@
 
-In the \emph{armed} scenario we modify the \emph{health} scenario by arming the survivors with pistols.
-Each armed agent has a pistol that fires 2 shots when activated.
-On firing, two rays are cast with a small random spread ($\angle p \sim \mathcal{N}(\angle aim, 2^\circ)$) for a range (2m) that covers most of the area.
+In the \emph{armed} scenario we attempt to rebalance the game by modifying the \emph{health} scenario to arm the survivors with firearms.
+Each armed agent has a firearm that fires 2 shots simultaneously when activated.
+On firing, two rays are cast with a small random spread ($\angle ray \sim \mathcal{N}(\angle aim, 2^\circ)$).
 When a ray first intersects with another agent it is counted as a `hit'.
@@
 
-We add a normalized current heading and relative heading of locations of other agents to an armed agent's state space, as well as a normalized time remaining until reloaded.
+We add a normalized current heading and relative heading of locations of other agents to an armed agent's state space, as well as a normalized time remaining until the firearm is reloaded.
+
 Reloading is automatic and does not require an action.
-The continuous Aiming action (left and right) modifies angular velocity of aim directly up to a maximum speed of 1 revolution per second.
-The firing mechanic is implemented as a `tirgger pressure' which determines the probability of a loaded pistol firing in a given timestep.
-Firing probability is implemented as the sigmoid function $p(fire) = \frac{1}{1+e^{20(x-0.5)}}$, where $x$ is the pressure on the trigger.
+The continuous aiming action (left and right) modifies angular velocity of aim directly up to a maximum speed of 1 revolution per second.
+The firing mechanic is implemented as a `trigger pressure' which determines the probability of a loaded firearm discharging in a given timestep.
+Firing probability is implemented as the sigmoid function $p(fire) = \frac{1}{1+e^{20(x-0.5)}}$, where $x$ is pressure on the trigger.
 
@@
 
+TODO: Mark
+
 \section{Discussion}
@@
 
-FROM HERE FOLLOWS TEMPLATE STUFF
-
-\paragraph{\LaTeX-specific details:}
-To use Times Roman in \LaTeX2e{}, put the following in the preamble:
-\begin{quote}
-\small
-\begin{verbatim}
-\usepackage{times}
-\usepackage{latexsym}
-\end{verbatim}
-\end{quote}
-
-\subsection{Ruler}
-A printed ruler (line numbers in the left and right margins of the article) should be presented in the version submitted for review, so that reviewers may comment on particular lines in the paper without circumlocution.
-The presence or absence of the ruler should not change the appearance of any other content on the page.
-The camera ready copy should not contain a ruler.
-
-\paragraph{Reviewers:}
-note that the ruler measurements may not align well with lines in the paper -- this turns out to be very difficult to do well when the paper contains many figures and equations, and, when done, looks ugly.
-In most cases one would expect that the approximate location will be adequate, although you can also use fractional references (\emph{e.g.}, this line ends at mark $295.5$).
-
-\paragraph{\LaTeX-specific details:}
-The style files will generate the ruler when {\small\verb|\aclfinalcopy|} is commented out, and remove it otherwise.
-
-\subsection{Title and Authors}
-\label{ssec:title-authors}
-
-Center the title, author's name(s) and affiliation(s) across both columns.
-Do not use footnotes for affiliations.
-Place the title centered at the top of the first page, in a 15-point bold font.
-Long titles should be typed on two lines without a blank line intervening.
-Put the title 2.5 cm from the top of the page, followed by a blank line, then the author's names(s), and the affiliation on the following line.
-Do not use only initials for given names (middle initials are allowed).
-Do not format surnames in all capitals (\emph{e.g.}, use ``Mitchell'' not ``MITCHELL'').
-Do not format title and section headings in all capitals except for proper names (such as ``BLEU'') that are
-conventionally in all capitals.
-The affiliation should contain the author's complete address, and if possible, an electronic mail address.
-
-The title, author names and addresses should be completely identical to those entered to the electronical paper submission website in order to maintain the consistency of author information among all publications of the conference.
-If they are different, the publication chairs may resolve the difference without consulting with you; so it is in your own interest to double-check that the information is consistent.
-
-Start the body of the first page 7.5 cm from the top of the page.
-\textbf{Even in the anonymous version of the paper, you should maintain space for names and addresses so that they will fit in the final (accepted) version.}
-
-\subsection{Abstract}
-Use two-column format when you begin the abstract.
-Type the abstract at the beginning of the first column.
-The width of the abstract text should be smaller than the
-width of the columns for the text in the body of the paper by 0.6 cm on each side.
-Center the word \textbf{Abstract} in a 12 point bold font above the body of the abstract.
-The abstract should be a concise summary of the general thesis and conclusions of the paper.
-It should be no longer than 200 words.
-The abstract text should be in 10 point font.
-
-\subsection{Text}
-Begin typing the main body of the text immediately after the abstract, observing the two-column format as shown in the present document.
-
-Indent 0.4 cm when starting a new paragraph.
-
-\subsection{Sections}
-
-Format section and subsection headings in the style shown on the present document.
-Use numbered sections (Arabic numerals) to facilitate cross references.
-Number subsections with the section number and the subsection number separated by a dot, in Arabic numerals.
-
-\subsection{Footnotes}
-Put footnotes at the bottom of the page and use 9 point font.
-They may be numbered or referred to by asterisks or other symbols.\footnote{This is how a footnote should appear.}
-Footnotes should be separated from the text by a line.\footnote{Note the line separating the footnotes from the text.}
-
-\subsection{Graphics}
-
-Place figures, tables, and photographs in the paper near where they are first discussed, rather than at the end, if possible.
-Wide illustrations may run across both columns.
-Color is allowed, but adhere to Section~\ref{ssec:accessibility}'s guidelines on accessibility.
-
-\paragraph{Captions:}
-Provide a caption for every illustration; number each one sequentially in the form:
-``Figure 1. Caption of the Figure.''
-``Table 1. Caption of the Table.''
-Type the captions of the figures and tables below the body, using 10 point text.
-Captions should be placed below illustrations.
-Captions that are one line are centered (see Table~\ref{font-table}).
-Captions longer than one line are left-aligned (see Table~\ref{tab:accents}).
-
-\begin{table}
-\centering
-\begin{tabular}{lc}
-\hline
-\textbf{Command} & \textbf{Output}\\
-\hline
-\verb|{\"a}| & {\"a} \\
-\verb|{\^e}| & {\^e} \\
-\verb|{\`i}| & {\`i} \\ 
-\verb|{\.I}| & {\.I} \\ 
-\verb|{\o}| & {\o} \\
-\verb|{\'u}| & {\'u}  \\ 
-\verb|{\aa}| & {\aa}  \\\hline
-\end{tabular}
-\begin{tabular}{lc}
-\hline
-\textbf{Command} & \textbf{Output}\\
-\hline
-\verb|{\c c}| & {\c c} \\ 
-\verb|{\u g}| & {\u g} \\ 
-\verb|{\l}| & {\l} \\ 
-\verb|{\~n}| & {\~n} \\ 
-\verb|{\H o}| & {\H o} \\ 
-\verb|{\v r}| & {\v r} \\ 
-\verb|{\ss}| & {\ss} \\
-\hline
-\end{tabular}
-\caption{Example commands for accented characters, to be used in, \emph{e.g.}, \BibTeX\ names.}\label{tab:accents}
-\end{table}
-
-\paragraph{\LaTeX-specific details:}
-The style files are compatible with the caption and subcaption packages; do not add optional arguments.
-\textbf{Do not override the default caption sizes.}
-
-\subsection{Hyperlinks}
-Within-document and external hyperlinks are indicated with Dark Blue text, Color Hex \#000099.
-
-\subsection{Citations}
-Citations within the text appear in parentheses as~\citep{Gusfield:97} or, if the author's name appears in the text itself, as \citet{Gusfield:97}.
-Append lowercase letters to the year in cases of ambiguities.  
-Treat double authors as in~\citep{Aho:72}, but write as in~\citep{Chandra:81} when more than two authors are involved. Collapse multiple citations as in~\citep{Gusfield:97,Aho:72}.
-
-Refrain from using full citations as sentence constituents.
-Instead of
-\begin{quote}
-  ``\citep{Gusfield:97} showed that ...''
-\end{quote}
-write
-\begin{quote}
-``\citet{Gusfield:97} showed that ...''
-\end{quote}
-
-\begin{table*}
-\centering
-\begin{tabular}{lll}
-\hline
-\textbf{Output} & \textbf{natbib command} & \textbf{Old ACL-style command}\\
-\hline
-\citep{Gusfield:97} & \small\verb|\citep| & \small\verb|\cite| \\
-\citealp{Gusfield:97} & \small\verb|\citealp| & no equivalent \\
-\citet{Gusfield:97} & \small\verb|\citet| & \small\verb|\newcite| \\
-\citeyearpar{Gusfield:97} & \small\verb|\citeyearpar| & \small\verb|\shortcite| \\
-\hline
-\end{tabular}
-\caption{\label{citation-guide}
-Citation commands supported by the style file.
-The style is based on the natbib package and supports all natbib citation commands.
-It also supports commands defined in previous ACL style files for compatibility.
-}
-\end{table*}
-
-\paragraph{\LaTeX-specific details:}
-Table~\ref{citation-guide} shows the syntax supported by the style files.
-We encourage you to use the natbib styles.
-You can use the command {\small\verb|\citet|} (cite in text) to get ``author (year)'' citations as in \citet{Gusfield:97}.
-You can use the command {\small\verb|\citep|} (cite in parentheses) to get ``(author, year)'' citations as in \citep{Gusfield:97}.
-You can use the command {\small\verb|\citealp|} (alternative cite without  parentheses) to get ``author year'' citations (which is useful for  using citations within parentheses, as in \citealp{Gusfield:97}).
-
-\subsection{References}
-Gather the full set of references together under the heading \textbf{References}; place the section before any Appendices. 
-Arrange the references alphabetically by first author, rather than by order of occurrence in the text.
-
-Provide as complete a citation as possible, using a consistent format, such as the one for \emph{Computational Linguistics\/} or the one in the  \emph{Publication Manual of the American 
-Psychological Association\/}~\citep{APA:83}.
-Use full names for authors, not just initials.
-
-Submissions should accurately reference prior and related work, including code and data.
-If a piece of prior work appeared in multiple venues, the version that appeared in a refereed, archival venue should be referenced.
-If multiple versions of a piece of prior work exist, the one used by the authors should be referenced.
-Authors should not rely on automated citation indices to provide accurate references for prior and related work.
-
-The following text cites various types of articles so that the references section of the present document will include them.
-\begin{itemize}
-\item Example article in journal: \citep{Ando2005}.
-\item Example article in proceedings, with location: \citep{borschinger-johnson-2011-particle}.
-\item Example article in proceedings, without location: \citep{andrew2007scalable}.
-\item Example arxiv paper: \citep{rasooli-tetrault-2015}. 
-\end{itemize}
-
-\paragraph{\LaTeX-specific details:}
-The \LaTeX{} and Bib\TeX{} style files provided roughly follow the American Psychological Association format.
-If your own bib file is named \texttt{\small acl2020.bib}, then placing the following before any appendices in your \LaTeX{}  file will generate the references section for you:
-\begin{quote}\small
-\verb|\bibliographystyle{acl_natbib}|\\
-\verb|\bibliography{acl2020}|
-\end{quote}
-
-You can obtain the complete ACL Anthology as a Bib\TeX\ file from \url{https://aclweb.org/anthology/anthology.bib.gz}.
-To include both the anthology and your own bib file, use the following instead of the above.
-\begin{quote}\small
-\verb|\bibliographystyle{acl_natbib}|\\
-\verb|\bibliography{anthology,acl2020}|
-\end{quote}
-
-\subsection{Digital Object Identifiers}
-As part of our work to make ACL materials more widely used and cited outside of our discipline, ACL has registered as a CrossRef member, as a registrant of Digital Object Identifiers (DOIs), the standard for registering permanent URNs for referencing scholarly materials.
-
-All camera-ready references are required to contain the appropriate DOIs (or as a second resort, the hyperlinked ACL Anthology Identifier) to all cited works.
-Appropriate records should be found for most materials in the current ACL Anthology at \url{http://aclanthology.info/}.
-As examples, we cite \citep{goodman-etal-2016-noise} to show you how papers with a DOI will appear in the bibliography.
-We cite \citep{harper-2014-learning} to show how papers without a DOI but with an ACL Anthology Identifier will appear in the bibliography.
-
-\paragraph{\LaTeX-specific details:}
-Please ensure that you use Bib\TeX\ records that contain DOI or URLs for any of the ACL materials that you reference.
-If the Bib\TeX{} file contains DOI fields, the paper title in the references section will appear as a hyperlink to the DOI, using the hyperref \LaTeX{} package.
-
-\subsection{Appendices}
-Appendices, if any, directly follow the text and the
-references (but only in the camera-ready; see Appendix~\ref{sec:appendix}).
-Letter them in sequence and provide an informative title:
-\textbf{Appendix A. Title of Appendix}.
-
-\section{Accessibility}
-\label{ssec:accessibility}
-
-In an effort to accommodate people who are color-blind (as well as those printing to paper), grayscale readability is strongly encouraged.
-Color is not forbidden, but authors should ensure that tables and figures do not rely solely on color to convey critical distinctions.
-A simple criterion:
-All curves and points in your figures should be clearly distinguishable without color.
-
-\section{Translation of non-English Terms}
-
-It is also advised to supplement non-English characters and terms with appropriate transliterations and/or translations since not all readers understand all such characters and terms.
-Inline transliteration or translation can be represented in the order of:
-\begin{center}
-\begin{tabular}{c}
-original-form \\
-transliteration \\
-``translation''
-\end{tabular}
-\end{center}
-
-\section{\LaTeX{} Compilation Issues}
-You may encounter the following error during compilation: 
-\begin{quote}
-{\small\verb|\pdfendlink|} ended up in different nesting level than {\small\verb|\pdfstartlink|}.
-\end{quote}
-This happens when \texttt{\small pdflatex} is used and a citation splits across a page boundary.
-To fix this, the style file contains a patch consisting of two lines:
-(1) {\small\verb|\RequirePackage{etoolbox}|} (line 455 in \texttt{\small acl2020.sty}), and
-(2) A long line below (line 456 in \texttt{\small acl2020.sty}).
-
-If you still encounter compilation issues even with the patch enabled, disable the patch by commenting the two lines, and then disable the \texttt{\small hyperref} package by loading the style file with the \texttt{\small nohyperref} option:
-
-\noindent
-{\small\verb|\usepackage[nohyperref]{acl2020}|}
-
-\noindent
-Then recompile, find the problematic citation, and rewrite the sentence containing the citation. (See, {\em e.g.}, \url{http://tug.org/errors.html})
-
-\section*{Acknowledgments}
-
-The acknowledgments should go immediately before the references. Do not number the acknowledgments section.
-Do not include this section when submitting your paper for review.
-
-\bibliography{anthology,acl2020}
+\bibliography{paper}
 \bibliographystyle{acl_natbib}
@@
 \appendix
+\section{Appendix}
+\label{sec:appendix}
 
-\section{Appendices}
-\label{sec:appendix}
-Appendices are material that can be read, and include lemmas, formulas, proofs, and tables that are not critical to the reading and understanding of the paper. 
-Appendices should be \textbf{uploaded as supplementary material} when submitting the paper for review.
-Upon acceptance, the appendices come after the references, as shown here.
-
-\paragraph{\LaTeX-specific details:}
-Use {\small\verb|\appendix|} before any appendix section to switch the section numbering over to letters.
-
-\section{Supplemental Material}
-\label{sec:supplemental}
-Submissions may include non-readable supplementary material used in the work and described in the paper.
-Any accompanying software and/or data should include licenses and documentation of research review as appropriate.
-Supplementary material may report preprocessing decisions, model parameters, and other details necessary for the replication of the experiments reported in the paper.
-Seemingly small preprocessing decisions can sometimes make a large difference in performance, so it is crucial to record such decisions to precisely characterize state-of-the-art methods.
-
-Nonetheless, supplementary material should be supplementary (rather than central) to the paper.
-\textbf{Submissions that misuse the supplementary material may be rejected without review.}
-Supplementary material may include explanations or details of proofs or derivations that do not fit into the paper, lists of
-features or feature templates, sample inputs and outputs for a system, pseudo-code or source code, and data.
-(Source code and data should be separate uploads, rather than part of the paper).
-
-The paper should not rely on the supplementary material: while the paper may refer to and cite the supplementary material and the supplementary material will be available to the reviewers, they will not be asked to review the supplementary material.
+TODO: Mark
 

### 2020-12-19 paper/paper.tex
update paper

@@
     Group submission for the final paper of XCS229ii,
-    \mbox{Stanford Center for Professional Development}
+    \mbox{Stanford Center for Professional Development}.
+    Code available at https://github.com/captainpete/zombie-survival.
   }
@@
 \begin{abstract}
-  The abstract goes here...
+  We investigate challenges of agent design using
+  multi-agent zero-sum game called ``Zombie-Survival'' 
+  in which we train 7 agents (2 survivors, 5 zombies)
+  under different scenarios using the same reward function
+  to investigate how different environmental dynamics
+  can lead to very different learned policies.
+
+  We show how state-space compression can improve the stability of a learned policy,
+  how a health mechanism that impacts maximum speed can change the dominant strategy,
+  and how arming the survivors with projectile weapons can lead to a social dilemma.
+
+  Ultimately the survivors do not prevail.
 \end{abstract}
@@
 
-Reinforcement Learning (RL) methods are seeing increasing use in real-world control systems.
-In part, the rising popularity of RL methods is due to its widely publicized success on challenging tasks,
-as well as the availability of quality open-source frameworks containing implementations of state-of-the-art models.
+Reinforcement Learning (RL) methods have enjoyed recent popularity,
+in part due to widely publicized successes on challenging tasks \citep{Silver1140,badia2020agent57,alphastarblog,barth2018distributed},
+and the availability of quality open-source frameworks containing implementations of state-of-the-art models \citep{1606.01540,hoffman2020acme,stable-baselines}.
 
-\subsection{Closing the Reality Gap}
+Though RL methods exceed human-level performance in some tasks such as
+simulated games, network optimisation \citep{chen2018auto,mao2019learning,valadarsky2017learning} and even flying helicopters \citep{abbeel2006application},
+there are many challenges discouraging the real-world deployment of RL systems \citep{dulacarnold2019challenges}.
 
-Current RL methods are not sample-effecicient, meaning the policy learner needs a lot of data before an effective policy can be found.
+\subsection{The Reality Gap}
+
+One of these challenges is that most RL methods are not sample-effecicient, meaning the policy learner needs a lot of data before an effective policy can be found.
 Often, there is not enough data available, or it cannot be captured safely (consider an autonomous driver learning the best policy for a near-crash scenario, where it will make mistakes in order to learn).
@@
 However, simulations often do not capture all the relevant real-world dynamics, so it can be difficult to transition a policy from the simulator to reality.
-This difficulty, known as the ``reality gap'', is often mitigated by adding noise to the simulation (to mimic noisy sensors and stochastic dynamics),
+This difficulty, known as the ``reality gap'', is often mitigated by adding noise \citep{jakobi1995noise} to the simulation (to mimic noisy sensors and stochastic dynamics),
 or by sufficiently improving the simulation to model the relevant dynamics.
-
-This approach has enjoyed success in single-agent RL settings like autonomous aircraft and robotics.
 
@@
 
-In cooperative settings like factories, or competitive settings like markets, using the simulation technique becomes more challenging;
-simulating the dynamics of other agents present in the environment is a difficult problem, even in simulation.
-This is especially the case when the policies of multiple agents are being modified and improved at the same time.
+When designing an RL agent for use in cooperative settings like factories, or competitive setting like markets,
+we have to allow for the fact that other agents may be present in the environment and may change their behavior in response to the activity of other agents.
+Training multiple agents concurrently in simulation can capture some of the dynamics around agents learning to respond to each other's actions.
+However, there are technical challenges to do with non-stationarity that prevent success in the naive approach of using multiple, single-agent algorithms.
 
-These issues in multi-agent settings motivated a training algorithm called Multi-Agent Deep Deterministic Policy Gradient (MADDPG) \citep{lowe2020multiagent}.
+These technical issues motivated the Multi-Agent Deep Deterministic Policy Gradient (MADDPG) algorithm \citep{lowe2020multiagent}.
 MADDPG addresses some issues in multi-agent learning, and allows running experiments that explore the emergence of competitive and cooperative policies in complex environments.
@@
 
-We explore MADDPG in the context of a mixed competitive-cooperative simulated environment called ``Zombie-Survival''.
-We make several modifications in order to test different scenarios.
-In these scenarios we show how specific kinds of environmental dynamics can affect changes in the evolution of policies.
+We explore the impacts of environmental dynamics on policy evolution 
+in the multi-agent RL setting.
+We use MADDPG to train multiple agents in a mixed competitive-cooperative simulated environment called ``Zombie-Survival''.
+We make several modifications to the environment in order to observe agent behavior under different scenarios.
 
-In (\ref{sec:background}) we review theory, RL algorithms, and aspects of agent design.
+In (\ref{sec:background}) we review theory, training algorithms, and aspects of agent design.
 In (\ref{sec:game}) we detail the basic simulation environment.
@@
 In (\ref{sec:health}) we alter the environment to cause a change in the value-function (keeping reward unchanged) to study the impact on learned policies.
-In (\ref{sec:arms}) we arm the survivors with projectile weapons to investigate how additional capability can make it harder for agents to achieve their objectives.
-We present methods and results for each scenario, and conclude with a discussion (\ref{sec:discussion}) of our findings.
+In (\ref{sec:arms}) we arm the survivors with projectile weapons to investigate how this additional capability can make it harder for agents to achieve their objectives.
+We present methods and results for each scenario, and conclude with a discussion of our findings (\ref{sec:discussion}).
 
@@
 
-A Markov Decision Process (MDP) formalizes the interaction of an agent with an environment.
+A Markov Decision Process (MDP) formalizes the interaction of an agent with an environment \citep{SuttonBarto}.
 At every time step, an \emph{agent} receives observations of the \emph{state} of the \emph{environment}, and sends \emph{actions} to the environment.
@@
 In the case when state transitions are not independent (i.e.
-the chance of arriving in a state depends on visiting previous states), then agent actions in a state can impact future expected rewards as well as immediate rewards.
+the chance of arriving in a state depends on visiting previous states), an agent's actions in a state can impact future expected rewards as well as immediate rewards.
 
 RL algorithms attempt to solve the MDP in order to maximize return (reward over time).
-This can be done by learning a policy (a function that maps from state to action).
+This is done by learning a policy (a function that maps from state to action).
 A policy can often be learned directly, or by learning a value function (a mapping from states and actions to expected returns) from which a policy can be extracted.
@@
 
-\paragraph{Deep Q-Learning Networks:} Q-Learning \citep{watkins} is an RL method that makes use of an action-value function $Q^\pi(s, a)$ corresponding to a policy $\pi$ (where $s$ is the state, and $a$ the action).
+\paragraph{DQN} Q-Learning \citep{watkins} is an RL method that makes use of an action-value function $Q^\pi(s, a)$ corresponding to a policy $\pi$ (where $s$ is the state, and $a$ the action).
 Deep Q-Learning Networks (DQN) \citep{2015Natur.518..529M} learn the action-value function $Q^*$ corresponding to the optimal policy by minimizing the loss function
@@
 
-\paragraph{Policy-gradient methods} learn a policy $\pi_\theta$, parameterized by $\theta$,
+\paragraph{Policy-gradient} Policy-gradient (PG) methods learn a policy $\pi_\theta$, parameterized by $\theta$,
 by sampling the gradient of a performance measure $J(\theta)$ with respect to $\theta$.
@@
 
-\paragraph{REINFORCE} \citep{Williams92simplestatistical}, a simple policy-gradient algorithm, uses all the steps in an episode to find $Q^\pi(s,a)=\mathbb{E}_\pi[G|s,a]$, where $G$ is the total reward of an episode.
+Learning a policy directly is useful in situations where extraction of a policy from a value function would be difficult (consider the scenario of a very large, or continuous action space).
+
+REINFORCE \citep{Williams92simplestatistical} is a simple Monte-Carlo policy-gradient algorithm that uses all the steps in an episode to sample $Q^\pi(s,a)=\mathbb{E}_\pi[G|s,a]$ (where $G$ is the expected return of an episode) to make a policy update.
 Because reward functions sometimes give high values for some states, this method can give high-variance gradient estimates.
@@
 
-\paragraph{Actor-critic} \citep{bhatnagar2009natural} improves on REINFORCE by learning a state-value function with baseline (critic), and an action-value function (actor) concurrently.
-Actor-critic also has the advantage that the parameters for both functions are updated at every step in an episode.
+\paragraph{Actor-critic} The Actor-Critic algorithm \citep{bhatnagar2009natural} improves on REINFORCE by learning a state-value function with baseline (critic), and an action-value function (actor) concurrently.
+Actor-Critic also has the advantage that the parameters for both functions are updated at every step in an episode.
 The policy gradient at each time step $t$ can be expressed:
@@
 
-\paragraph{Deterministic Policy Gradient (DPG)} \citep{pmlr-v32-silver14} algorithms extend policy gradient methods to deterministic policies.
+\paragraph{DPG} The Deterministic Policy Gradient (DPG) algorithm \citep{pmlr-v32-silver14} extend policy-gradient methods to deterministic policies.
 A deterministic policy ($\mu_\theta$) maps a state to an action,
@@
 
-DPG methods are often used in continuous control tasks
+DPG methods are suited to continuous control tasks
 (taking the gradient of $Q^\mu$ with respect to $a$ requires the action space to be continuous).
 
-\paragraph{Deep Deterministic Policy Gradient (DDPG)} \citep{lillicrap2019continuous} combines ideas from DPG and DQN.
+\paragraph{DDPG} The Deep Deterministric Policy-Gradient algorithm \citep{lillicrap2019continuous} combines ideas from DPG and DQN.
 In DDPG the actor and critic are approximated with neural networks.
@@
 
-The Markov Decision Process (MDP) framework regards other agents as part of the environment from the perspective of a single agent.
-From the perspective of an agent, the influence on rewards from other agents in the environment can make it harder to learn a policy;
+Under MDP additional agents are regarded as part of the environment.
+From the perspective of an agent, the influence on rewards from other agents can make it harder to learn a policy;
 this is especially so when the environment contains other agents that are also learning policies,
-because elements of the `environment' change while the agent is trying to learn them.
+because the other agents can cause the environment's dynamics to change while the agent is trying to learn them.
 
-\citealp{lowe2020multiagent} showed that the chance of sampling a policy-gradient update in an improved direction decreases exponentially with the number of agents being trained in the shared environemnt.
+\citealp{lowe2020multiagent} showed that the chance of sampling a policy-gradient update for an agent in the direction of higher expected return decreases exponentially with the number of agents being trained in the shared environemnt.
 
@@
 In MADDPG, the critic $Q^\pi_i(x, a_1, a_2, ..., a_N)$ for agent $i$ has access to the actions of all $N$ agents, and some state information $x$.
-The critic is often only used in training, so it's possible to deploy MADDPG actors in continuous control tasks where knowledge of other actors' actions are not available.
-Agents can have different state and action spaces, and be goverened by different reward functions.
+%The critic is often only used in training, so it's possible to deploy MADDPG actors in continuous control tasks where knowledge of other actors' actions are not available.
+
+Under MADDPG, agents agents are permitted to have different state and action spaces, and to be goverened by different reward functions.
 We use the MADDPG algorithm for all agents within our scenarios.
@@
 
-When implementing an RL system there are many important design considerations:
+When implementing an RL system there are many important design decisions, some key considerations are:
 
@@
   \item \emph{Action-space}: The action-space describes actions available to the agent.
-The level of abstraction here should be fine enough for precise control, but general enough for a policy to be learnable.
+The level of abstraction here should be fine enough for precise control, but general enough for a policy to be easily learnable.
   \item \emph{Reward function}: The reward function provides a scalar feedback value (positive or negative) to the agent.
@@
 Crafting a consistent reward function that reflects the exact desires of the designer can be challenging \citep{leike2017ai}.
-Some algorithms can infer the a reward function from feedback \citep{ng2000algorithms}, but most simply try to maximize the explicit reward from a hard-coded function.
+Some algorithms can infer the a reward function from feedback \citep{ng2000algorithms}, but most simply try to maximize the explicit reward given by the function.
 \end{itemize}
 
-In our scenarios we examine the imapct of design choices for state-space and action-space under a fixed reward-function.
+In our scenarios we implement a simple reward function that creates a zero-sum game, where one agent's loss is another's gain.
+Our investigation examines the effect of design choices relating to the state-space and action-space, and how those impact the emergence of policies.
 
@@
 
-Zombies have the advantage of numbers (there are 5 zombies and 2 survivors), whereas survivors have the advantage of faster accleration and a higher maximum speed.
-All reward functions are independent, agents do not share reward.
-The game offers a scenario with asymmetric power dynamics between competitive teams, and allows the emergence of competitive or cooperative behavior within a team.
+Zombies have the advantage of numbers, whereas survivors have the advantage of faster accleration and a higher maximum speed (Table \ref{game-settings}).
 
@@
 Each agent is represented as a circle that has mass, position, velocity, velocity dampening, maximum acceleration, and maximum speed.
-The predator-prey environment uses reward shaping to discourage agents from moving out of bounds.
-We remove this additional term in the reward function and implement fixed boundary walls.
-We implement the biting rules, modify the agent count and adjust the agents' physical capabilities (see Table \ref{game-settings}).
-We run several training jobs for 10,000 steps each to capture the emergence of different policies.
+The predator-prey environment uses reward shaping to discourage agents from moving out of bounds; we remove this additional term in the reward function and implement fixed boundary walls.
+We implement the biting rules, modify the agent count and adjust the agents' physical capabilities (Table \ref{game-settings}).
+We run multiple experiments in which we train agents over 10,000 episodes.
 
@@
   \includegraphics[width=0.5\textwidth]{figures/baseline_reward.png}
-  \caption{Reward (moving average over 100 episodes) for survivors (green) and zombies (red) for 10 training runs.}
+  \caption{
+    Reward (moving average over 100 episodes) for survivors (green) and zombies (red) for 10 training runs.
+    Agents experience policy shifts between episode 2000 to 6000.
+    Policies stabilize after 7000 episodes.
+  }
   \label{fig:baseline_reward}
@@
 
-\paragraph{Variability:} We observed variability in policy convergence between training runs.
+\paragraph{Variability} We observed variability in policy convergence between training runs.
 Figure \ref{fig:baseline_reward} shows agent reward graphs over multiple runs.
@@
   \includegraphics[width=0.5\textwidth]{figures/baseline_speed.png}
-  \caption{Reward and speed (moving averages over 200 episodes) for a single training run in baseline scenario. Survivors in blue and orange.}
+  \caption{
+    Reward and average speed (moving averages over 200 episodes)
+    for all agents in a single training run of the \emph{baseline} scenario.
+    Survivors (blue and orange) learn faster movement at different times.
+  }
   \label{fig:baseline_speed}
@@
 
-\paragraph{Target fixation:} On further investigation we discovered the reason for the variability; a zombie usually fixates on a target, but zombies learn policies at different rates.
+\paragraph{Target fixation} On further investigation we discovered the reason for the variability; a zombie usually fixates on a target, but zombies learn policies at different rates.
 Figure \ref{fig:baseline_speed} shows speed and reward over time for a single training run.
@@
 
-In the baseline state-space, every agent is unique, so all agents have to learn their expected reward relating to each other.
-From an agent's perspective, when considering reward the most important thing about an agent is whether they are a survivor or a zombie, not necessarily which survivor or zombie.
-We'd prefer agents learn general policies that consider other agents exchangeable within a team.
+In the \emph{baseline} state-space, every agent is unique, so all agents have to learn the expected reward specific to every other agent.
+We propose that from an agent's perspective, when considering reward the most important thing about another agent is whether they are a survivor or a zombie, not necessarily which survivor or zombie.
+We'd prefer agents learn general policies that consider other agents as exchangeable within a team.
 
@@
 
-We expect both survivors and zombies to learn more general policies with faster convergence than the baseline.
+We expect both survivors and zombies to learn more general policies with faster convergence than in the \emph{baseline} scenario.
 We also expect to see zombies chasing which ever survivor is easiest to chase, rather than fixating on a single survivor.
@@
 
-We observed faster convergence to a stable policy (as judged by average speed and reward) when compared to baseline (see Figure \ref{fig:anon_base_compare}).
+We observed less variance in convergence to a stable policy (as judged by average speed and reward) when compared to \emph{baseline} (see Figure \ref{fig:anon_base_compare}).
 We also observed more uniform persuit and biting (see Figure \ref{fig:anon_base_dist_bite}).
@@
   \caption{
-    Speed and reward evolution for baseline and anonymous scenarios.
+    Speed and reward evolution for baseline and anonymous scenarios over 8 training runs.
     Results taken as moving averages over 200 episodes prior to aggregation.
@@
   \caption{
-    Distance and bite count comparison between baseline and anonymous scenarios.
-    Plot colors indicate average distance and average bite count per episode average across the last 5000 episodes of a training run.
-    Both axes are agent indexes, and color indicates relative value with white being the highest value.
+    Distance and bite count comparison between \emph{baseline} and \emph{anonymous} scenarios.
+    Plot colors indicate average distance and average bite count per episode, averaged across the last 5000 episodes of a single training run.
+    Both axes are agent indexes, and shade of color indicates relative value with white being the highest value.
   }
@@
 In the \emph{health} scenario we modify the \emph{anonymity} scenario to introduce a health mechanism for survivors.
-Survivors start each episode with 100\% health, which for survivors is discounted ($\gamma=0.99$) every bite.
+Survivors start each episode with 100\% health, which for survivors is discounted ($\gamma=0.99$) every bite (one bite per zombie in contact per time step).
 An agent's health level is added to the private observation space for the agent.
-Reward remains the same for bites regardless of health levels,
-however maximum speed at each step is limited to the fraction of health remaining.
+Maximum speed at each step is limited to the fraction of health remaining.
 If a survivor is bitten too often, they become slower than the zombies and risk being overrun, leading to a large amount of bites for the remainder of the episode.
@@
 
-\paragraph{Breakout:} Compared to the \emph{anonymous} scenario, agents in the \emph{health} scenario were slower on average (see Figure \ref{fig:anon_health_compare}, plot 1).
-It's interesting to note this slowdown occurred before the reward ramp-up due to zombies successfully immobilizing survivors.
-In part the slower speeds can be explained through partial health reductions, but rendering the simulation frames provided another explanation.
+\paragraph{Breakout} Compared to the \emph{anonymous} scenario, agents in the \emph{health} scenario were slower on average (see Figure \ref{fig:anon_health_compare}, plot 1).
+This slowdown occurred before the reward ramp-up caused by zombies learning to immobilize survivors.
+In part the initial slower speeds can be explained through partial health reductions, but rendering the simulation frames provided another explanation.
 In previous scenarios, survivors would run into walls at maximum speed to bounce back through the persuring group of zombies without losing speed.
-The bouncing strategy sacrifices minimal reward and allows a survivor to keep their distance ahead of the pursuing zombies after the manouver.
-In the \emph{health} scenario, this strategy becomes unsustainable, as every contact with a zombie (during the bounce) reduces the survivors maximum speed
-leading to a very bad outcome after only a few bounces.
+The bouncing strategy sacrifices reward but allows a survivor to keep their distance ahead of the pursuing zombies after the manouver.
+In the \emph{health} scenario, this strategy becomes unsustainable, as every contact with a zombie reduces the survivors maximum speed leading to a very bad outcome after only a few bounces.
 
-With health to consider, survivors opted for more impressive fake-and-dodge manouvers to avoid contact altogether.
-Often this would involve pausing to allow zombies to move closer to their target, then making use of the survivors' higher acceleration to escape at an angle orthagonal to the zombie group's velocity.
+\paragraph{Game-over} We measured a drastic increase in average zombie reward (Figure \ref{fig:anon_health_compare}, plot 2) due to zombies learning to immobilize and bite survivors.
+In most cases, one survivor would succumb to the zombies first.
+Zombies learn that moving in the direction of the slower-moving survivor leads to states with higher expected value, so they collaborate in slowing down one survivor, leaving the other survivor to escape out of harm's way (Figure \ref{fig:anon_health_compare}, plot 4).
 
-\paragraph{Game-over:} We measured a drastic increase in average zombie reward (see Figure \ref{fig:anon_health_compare}, plot 2) due to zombies learning to immobilize and bite survivors.
-In most cases, one survivor would succumb to the zombies first.
-Zombies learn that moving in the direction of the slower-moving survivor leads to states with higher expected value, so they collaborate in slowing down one survivor, leaving the other survivor to escape out of harm's way.
+\paragraph{Survivor competition} Survivors that were not the focus of `zombie collaboration' learned to stay out of the way by distancing themselves from the other survivor and their zombie entourage (see Figure \ref{fig:health_survivor_dist}).
 
-%We found it difficult to predict at the beginning of an episode which survivor would ultimately attract the zombies.
-%In watching many episodes we found a survivor having many zombies in tow would move directly toward the other survivor (who was often be keeping their distance in a corner).
-%The first survivor would have higher velocity, and use it to risk a wall-bounce manouver near the second slower-moving one.
-%The first survivor in this instance would retain the speed required to escape while the zombies switched their persuit to overrun the second survivor.
+In some episodes, survivors would trade places as the focus of zombie attention.
+A fast moving survivor with zombies in persuit would move directly toward the other survivor (who was often be keeping their distance in a corner).
+The first survivor would have higher velocity, and use it to risk a wall-bounce manouver near the second slower-moving one.
+The first survivor in this instance would retain the speed required to escape while the zombies switched their persuit to overrun the second survivor.
+We found it difficult initially to predict which survivor would ultimately attract the zombies, but around episode 6000 the survivors started to better avoid this trap by moving away from areas that would soon receive zombies.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{figures/health_survivor_dist.png}
+  \caption{
+    Distance between survivors in the \emph{health} scenario.
+    Survivors learn to keep their distance from one another more than in the \emph{anonymous} scenario.
+    Results taken as moving averages over 200 episodes prior to aggregation.
+  }
+  \label{fig:health_survivor_dist}
+\end{figure}
+
+The introduction of the health dynamic moved the simulation closer to what we consider a `more realistic' zombie-survivor scenario.
+The change in state-space and environmental dynamics lead to a very different movement policy for agents without changing the reward function.
 
@@
   \caption{
-    Comparison between anonymous and health scenario policy evolution.
+    Comparison between \emph{anonymous} and \emph{health} scenario policy evolution.
     Results taken as moving averages over 200 episodes prior to aggregation.
@@
 
-We add a normalized current heading and relative heading of locations of other agents to an armed agent's state space, as well as a normalized time remaining until the firearm is reloaded.
+We add a normalized current heading and relative heading of locations of other agents to a each survivor's state-space, as well as a normalized time remaining until the firearm is reloaded.
+Zombies and the other survivor do not observe a survivor's aim or reloading states.
 
 Reloading is automatic and does not require an action.
-The continuous aiming action (left and right) modifies angular velocity of aim directly up to a maximum speed of 1 revolution per second.
+The continuous aiming action (left and right) modifies angular velocity of aim directly (rather than through acceleration) up to a maximum speed of 1 revolution per second.
 The firing mechanic is implemented as a `trigger pressure' which determines the probability of a loaded firearm discharging in a given timestep.
-Firing probability is implemented as the sigmoid function $p(fire) = \frac{1}{1+e^{20(x-0.5)}}$, where $x$ is pressure on the trigger.
+Firing probability is implemented as the sigmoid function $p(fire) = \frac{health}{1+e^{20(x-0.5)}}$, where $x$ is pressure on the trigger (note the ability for a survivor to fire depends on their health).
 
@@
 
-TODO: Mark
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{figures/armed_health_compare.png}
+  \caption{
+    Comparison between \emph{armed} and \emph{health} scenario policy evolution over 3 training runs.
+    Results taken as moving averages over 500 episodes prior to aggregation.
+    Plot 1: average speed zombies and survivors per episode.
+    Plot 2: survivor episode reward.
+    Plot 3: episode-end health of zombies and survivors.
+    Plot 4: total hits recieved by zombies and survivors.
+  }
+  \label{fig:armed_health_compare}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{figures/survivor_skill.png}
+  \caption{
+    Comparison of hits from different survivors in the \emph{armed} scenario over 3 training runs.
+    Results taken as moving averages over 1000 episodes prior to aggregation due to high variance.
+    Plot 1: zombie hit counts for each survivor.
+    Plot 2: survivor hit counts for each survivor.
+    Plot 3: shots fired for each survivor.
+    Plot 4: ratio of zombie and survivor hit counts to shots fired for each survivor.
+    Plot 5: probability a survivor shot the other survivor, conditional on them having fired and hit something, normaled for team size.
+  }
+  \label{fig:survivor_skill}
+\end{figure}
+
+As expected, the introduction of projectile weapons in the \emph{armed} scenario increased survivor health and rewards on average (Figure \ref{fig:armed_health_compare}, plots 2-3),
+shifting the balance of power back toward the survivors.
+Relative to the \emph{health} scenario, survivors took longer to learn to increase average speed (Figure \ref{fig:armed_health_compare}, plot 1).
+This can be explained by the introduction of a mechanicm for reducing zombie health (the guns), leading to lower average maximum speeds for zombies,
+decreasing the urgency with which survivors must learn to move faster.
+
+Interestingly, the initial period (episodes 2000-4000) that survivors in the \emph{health} scenario used to learn increased speed
+is used by survivors in the \emph{armed} scenario to learn increased fire rate.
+
+\paragraph{Specialization}
+During this phase we found one survivor was often faster than the other at learning to hit targets (Figure \ref{fig:survivor_skill}, plot 1).
+The difference in rate of fire in survivors persists over a period and may be indicative of role specialization,
+where one survivor learns a benefit from shooting while the other finds more benefit in learning evasion.
+
+\paragraph{More practice required}
+After the initial phase, average survivor speed increased to a similar level to that observed in the \emph{health} scenario.
+Survivors were able to hold back the zombies initially, but the zombie's power-in-numbers eventually overcame the survivor's advantage.
+Examining the zombie hit count compared to the survivor shots fired (Figure \ref{fig:survivor_skill}, plots 1, 3)
+indicates that survivors did not become sharpshooters within the number of simulated episodes.
+We think this could be improved slightly by simplifying the firing action, or by increasing the incentive (indirectly) to learn accuracy with more powerful guns and a longer reload time.
+Or potentially using transfer learning from a `firing range' scenario.
+
+\paragraph{Social dilemmas}
+We observed the emergence of another competitive strategy among survivors.
+In some episodes we noticed a survivor would occasionally shoot another survivor.
+The slowdown caused to the target by the shot would sometimes cause them to become overrun with zombies,
+leaving the shooter free of zombie aggravation.
+
+To establish whether this was a strategy and not just chance
+we modelled the conditional probability of defection (shooting another survivor) as a Bernoulli trial
+in which the survivor, having fired and hit something, has either hit the other survivor or a zombie.
+
+\begin{equation}
+  p_i(defect|hit, i\in S) = \frac{
+    \sum\limits_{j\neq i, j\in S} h_{i,j}
+  }{
+    \sum\limits_{j\neq i, j\in S} h_{i,j}
+    +
+    \sum\limits_{j\in Z} h_{i,j}
+  }
+\end{equation}
+
+where $h_{i,j}$ is the number of times agent $i$ shot agent $j$ in an episode, $S$ is the set of survivors and $Z$ the set of zombies.
+
+We found that if we accept the model then survivors are learning that if they can hit an agent, a survivor is higher value target than a zombie (Figure \ref{fig:survivor_skill}, plot 5).
+We were unable to train the model for more than 20,000 episodes (due to time constraints) to determine if defection emerges as a dominant strategy over time.
 
@@
 
+In attempting to design scenarios to model a `real-life' zombie-survival situation we witnessed the emergence of surprising behaviors.
+In each case, our surprise was a result of the environment's dynamics (as implemented) not matching our expectations.
+
+In the \emph{baseline} scenario, we didn't expect target-fixation because our personal state-space representation was already abstracting the agents into two teams.
+When we aligned the state-space using our personal abstractions in the \emph{anonymous} scenario the behavior was corrected, but we encountered further surprises.
+We did not expect the wall-bouncing strategy because our personal expectation was that running into a wall when chased by zombies is counterproductive.
+
+Our explanation for this assumption was to do with health, which we weren't modelling.
+Once we corrected this in the \emph{health} scenario the strategy became less common and we observed the behavior we expected.
+
+In the \emph{armed} scenario we weren't certain how agents would respond.
+Survivors were less effective than expected within the training time at using weapons, likely because we underestimated the complexities of a task achievable by humans (in popular depictions of zombie-survivor scenarios, survivors are adept at shooting).
+
+The competitive behavior we saw between survivors was also a surprise, as we expected the survivors to learn the advantage of cooperation.
+On reflection, that expectation was a result of our personal ethical conditioning and our anthropomorphizing the survivors.
+A more thorough empirical game theoretic analysis such as performed in \citealp{10.5555/3091125.3091194} would likely help us understand the agent incentives.
+
+We found it an interesting challenge to attempt to produce desired behavior in a multi-agent setting.
+Above all, we learned that agents will seek to maximize their reward functions, rather than do what you expect them to do.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{figures/screencapture.png}
+  \caption{
+    The \emph{armed} environment.
+    Survivors in green, zombies in red.
+    Both survivors are firing weapons.
+  }
+  \label{fig:baseline_reward}
+\end{figure}
+
 \bibliography{paper}
@@
 
-\appendix
-\section{Appendix}
-\label{sec:appendix}
-
-TODO: Mark
-
 \end{document}

### 2020-12-20 paper/paper.tex
update paper

@@
 
-\subsection{Agent Design}
-
-When implementing an RL system there are many important design decisions, some key considerations are:
-
-\begin{itemize}
-  \item \emph{Algorithm}: Different algorithms are suited to different tasks.
+% \subsection{Agent Design}
+% 
+% When implementing an RL system there are many important design decisions, some key considerations are:
+% 
+% \begin{itemize}
+%   \item \emph{Algorithm}: Different algorithms are suited to different tasks.
 A selected algorithm will need to perform well with the selected action-space, state-space, reward function and environmental dynamics.
-  \item \emph{State-space}: The state-space describes how environmental observations are presented to the algorithm.
+%   \item \emph{State-space}: The state-space describes how environmental observations are presented to the algorithm.
 Constructing the correct, relevant abstractions allow an agent to better learn the environment.
-  \item \emph{Action-space}: The action-space describes actions available to the agent.
+%   \item \emph{Action-space}: The action-space describes actions available to the agent.
 The level of abstraction here should be fine enough for precise control, but general enough for a policy to be easily learnable.
-  \item \emph{Reward function}: The reward function provides a scalar feedback value (positive or negative) to the agent.
+%   \item \emph{Reward function}: The reward function provides a scalar feedback value (positive or negative) to the agent.
 Some algorithms can model uncertainty in the reward function \citep{ghavamzadeh2016bayesian}, but most expect a point-estimate rather than a distribution.
@@
 Some algorithms can infer the a reward function from feedback \citep{ng2000algorithms}, but most simply try to maximize the explicit reward given by the function.
-\end{itemize}
-
-In our scenarios we implement a simple reward function that creates a zero-sum game, where one agent's loss is another's gain.
-Our investigation examines the effect of design choices relating to the state-space and action-space, and how those impact the emergence of policies.
+% \end{itemize}
+% 
+% In our scenarios we implement a simple reward function that creates a zero-sum game, where one agent's loss is another's gain.
+% Our investigation examines the effect of design choices relating to the state-space and action-space, and how those impact the emergence of policies.
 
@@
 Survivors were less effective than expected within the training time at using weapons, likely because we underestimated the complexities of a task achievable by humans (in popular depictions of zombie-survivor scenarios, survivors are adept at shooting).
-
-The competitive behavior we saw between survivors was also a surprise, as we expected the survivors to learn the advantage of cooperation.
-On reflection, that expectation was a result of our personal ethical conditioning and our anthropomorphizing the survivors.
-A more thorough empirical game theoretic analysis such as performed in \citealp{10.5555/3091125.3091194} would likely help us understand the agent incentives.
+The social dilemma presented warrants further study;
+a more thorough empirical game theoretic analysis such as performed in \citealp{10.5555/3091125.3091194} would be an interesting exercise.
 

### 2020-12-20 paper/paper.tex
update paper and schenario code comments

@@
 
-We found that if we accept the model then survivors are learning that if they can hit an agent, a survivor is higher value target than a zombie (Figure \ref{fig:survivor_skill}, plot 5).
+We found that if we accept the model then survivors are learning that if they can hit an agent, it is more valuable to hit a survivor (Figure \ref{fig:survivor_skill}, plot 5).
 We were unable to train the model for more than 20,000 episodes (due to time constraints) to determine if defection emerges as a dominant strategy over time.
@@
 
-In attempting to design scenarios to model a `real-life' zombie-survival situation we witnessed the emergence of surprising behaviors.
-In each case, our surprise was a result of the environment's dynamics (as implemented) not matching our expectations.
+In attempting to design scenarios to model a `real-life' zombie-survival situation we encountered unexpected agent policies.
+We did not expect target-fixation, wall-bouncing or survivors having trouble using weapons.
 
-In the \emph{baseline} scenario, we didn't expect target-fixation because our personal state-space representation was already abstracting the agents into two teams.
-When we aligned the state-space using our personal abstractions in the \emph{anonymous} scenario the behavior was corrected, but we encountered further surprises.
-We did not expect the wall-bouncing strategy because our personal expectation was that running into a wall when chased by zombies is counterproductive.
+When we encoded additional dynamics into the environment to make our expectations explicit we caused changes in agent policy; only some of these changes were successful.
+Altering an environment and modifying the state-space as we have done is one way to elicit desired policies,
+another approach is modifying the reward function to add detail and caveats to the agent's high-level goal.
+Both methods look to be blunt instruments for a real-world setting.
 
-Our explanation for this assumption was to do with health, which we weren't modelling.
-Once we corrected this in the \emph{health} scenario the strategy became less common and we observed the behavior we expected.
+An implication of the challenges we saw is that if an RL agent were to be deployed in a real-world environment,
+the agent's behavior may only be consistent with expectations to the extent that the environment dynamics remain stationary.
+Beyond the known environmental dynamics, it is hard to know ahead of time (without modelling and simulation) what agent behavior would result.
+Additionally, policies learned are not Bayes-optimal, so an agent may not always learn a desired policy even in the presence of a well expressed reward function and a perfectly representative simulation environment.
 
-In the \emph{armed} scenario we weren't certain how agents would respond.
-Survivors were less effective than expected within the training time at using weapons, likely because we underestimated the complexities of a task achievable by humans (in popular depictions of zombie-survivor scenarios, survivors are adept at shooting).
-The social dilemma presented warrants further study;
-a more thorough empirical game theoretic analysis such as performed in \citealp{10.5555/3091125.3091194} would be an interesting exercise.
+One can make the argument that all real-world environments are multi-agent settings,
+so the study of multi-agent learning dynamics is especially interesting for real-world applications.
+
+A potential avenue for further exploration of the Zombie-Survival game
+would be a more thorough, empirical game-theoretic analysis, such as undertaken in \citealp{10.5555/3091125.3091194}.
 
 We found it an interesting challenge to attempt to produce desired behavior in a multi-agent setting.
-Above all, we learned that agents will seek to maximize their reward functions, rather than do what you expect them to do.
+While we had some success, we would hesitate before deploying our agents into the real world.
 

### 2020-12-20 multiagent/scenarios/zombie_health.py

@@ -8,10 +8,10 @@ from multiagent.scenario import BaseScenario
 #   2D continuous action space (up/down, left/right)
 
 # observation space:
-#   Complete information on 2D positions of all agents (continuous 4D *
+#   Partial information on 2D positions of all agents (continuous 4D *
 #   agents). This includes absolute position and velocity for the agent itself,
 #   and relative position and velociry for other humans, and other zombies
-#   (agents have identity by index position).
+#   (agents' positions are sorted by team then by distance to the observer).
 
 # reward function:
 #   When a zombie and human come into contact then 10 reward is deducted from
